### 2024-10-22 paper/paper.qmd
uploaded first drafts of all scripts, models, and visualizations for paper

---
title: "Modeling the 2024 US Presidential Elections"
subtitle: "My subtitle if needed"
author: 
  - Sophia Brothers
  - Deyi Kong
thanks: "Code and data are available at: [https://github.com/eeeee-cmd/US_Election/](https://github.com/eeeee-cmd/US_Election/)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

# Introduction

Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.

# Data

## Overview

This data was analyzed using [@citeR]. We also used the packages: tidyverse [@tidyverse], etc

## Measurement

## Outcome Variables

## Predictor Variables

# Model

## Model Set-Up

## Model Justification

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
library(tidyverse)
library(dplyr)
library(here)

cleaned_data <- read_csv(here::here("data/analysis_data/model_data.csv"))
lm_model <- readRDS(here::here("models/lm_model.rds"))

cleaned_data <- cleaned_data %>%
  mutate(PredictedPercentage = predict(lm_model, newdata = cleaned_data))

results <- cleaned_data %>%
  group_by(CandidateName) %>%
  summarise(AveragePredictedPercentage = mean(PredictedPercentage, na.rm = TRUE)) %>%
  arrange(desc(AveragePredictedPercentage)) %>%
  filter(AveragePredictedPercentage > 0)

total_percentage <- sum(results$AveragePredictedPercentage, na.rm = TRUE)

results <- results %>%
  mutate(NormalizedPercentage = (AveragePredictedPercentage / total_percentage) * 100)

print(results)
```

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(tidyverse)
library(here)

cleaned_data <- read_csv(here::here("data/analysis_data/electoral_data.csv"))
lm_model_trump <- readRDS(here::here("models/lm_model_trump.rds"))
lm_model_harris <- readRDS(here::here("models/lm_model_harris.rds"))

unique_states <- distinct(cleaned_data, State)

prediction_data <- cleaned_data %>%
  group_by(State) %>%
  summarise(
    PollScore = mean(PollScore, na.rm = TRUE),
    SampleSize = mean(SampleSize, na.rm = TRUE),
    NumericGrade = mean(NumericGrade, na.rm = TRUE),
    PollRecency = mean(PollRecency, na.rm = TRUE),
    .groups = 'drop'
  )

prediction_data <- prediction_data %>%
  mutate(
    TrumpPredicted = predict(lm_model_trump, newdata = prediction_data),
    HarrisPredicted = predict(lm_model_harris, newdata = prediction_data)
  )

electoral_votes <- tibble(
  State = c("Alabama", "Kentucky", "North Dakota", "Alaska", "Louisiana", 
            "Ohio", "Arizona", "Maine", "Oklahoma", "Arkansas", 
            "Maryland", "Oregon", "California", "Massachusetts", "Pennsylvania", 
            "Colorado", "Michigan", "Rhode Island", "Connecticut", "Minnesota", 
            "South Carolina", "Delaware", "Mississippi", "South Dakota", 
            "District of Columbia", "Missouri", "Tennessee", "Florida", 
            "Montana", "Texas", "Georgia", "Nebraska", "Utah", 
            "Hawaii", "Nevada", "Vermont", "Idaho", "New Hampshire", 
            "Virginia", "Illinois", "New Jersey", "Washington", 
            "Indiana", "New Mexico", "West Virginia", "Iowa", 
            "New York", "Wisconsin", "Kansas", "North Carolina", 
            "Wyoming"),
  Votes = c(9, 8, 3, 3, 8, 
            17, 11, 4, 7, 6, 
            10, 8, 54, 11, 19, 
            10, 15, 4, 7, 10, 
            9, 3, 6, 3, 
            3, 10, 11, 30, 
            4, 40, 16, 5, 6, 
            4, 6, 3, 4, 4, 
            13, 19, 14, 12, 
            11, 5, 4, 6, 
            28, 10, 6, 16, 
            3)
)

predictions_with_votes <- prediction_data %>%
  left_join(electoral_votes, by = "State")

predictions_with_votes <- predictions_with_votes %>%
  mutate(
    Winner = case_when(
      TrumpPredicted > HarrisPredicted ~ "Donald Trump",
      HarrisPredicted > TrumpPredicted ~ "Kamala Harris",
      TRUE ~ "Tie"
    ),
    TrumpElectoralVotes = if_else(Winner == "Donald Trump", Votes, 0),
    HarrisElectoralVotes = if_else(Winner == "Kamala Harris", Votes, 0)
  )

electoral_totals <- predictions_with_votes %>%
  summarise(
    TrumpTotalElectoralVotes = sum(TrumpElectoralVotes, na.rm = TRUE),
    HarrisTotalElectoralVotes = sum(HarrisElectoralVotes, na.rm = TRUE)
  )

print(electoral_totals)

state_winners <- predictions_with_votes %>%
  select(State, TrumpPredicted, HarrisPredicted) %>%
  mutate(
    Winner = case_when(
      TrumpPredicted > HarrisPredicted ~ "Donald Trump",
      HarrisPredicted > TrumpPredicted ~ "Kamala Harris",
      TRUE ~ "Tie"
    )
  )

print(state_winners, n=54)

```

# Results

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv"))

summary_stats <- cleaned_data %>%
  summarise(
    AveragePollScore = mean(PollScore, na.rm = TRUE),
    AverageSampleSize = mean(SampleSize, na.rm = TRUE),
    AveragePercentage = mean(Percentage, na.rm = TRUE),
    TotalPolls = n()
  )

print(summary_stats)

ggplot(cleaned_data, aes(x = PollScore)) +
  geom_histogram(bins = 30, fill = "blue", alpha = 0.7) +
  labs(title = "Distribution of Poll Scores", x = "Poll Score", y = "Frequency")

ggplot(cleaned_data, aes(x = SampleSize)) +
  geom_histogram(bins = 30, fill = "green", alpha = 0.7) +
  labs(title = "Distribution of Sample Size", x = "Sample Size", y = "Frequency")

ggplot(cleaned_data, aes(x = Percentage)) +
  geom_histogram(bins = 30, fill = "orange", alpha = 0.7) +
  labs(title = "Distribution of Percentage", x = "Percentage", y = "Frequency")

ggplot(cleaned_data, aes(x = CandidateName, y = Percentage)) +
  geom_boxplot() +
  labs(title = "Percentage by Candidate", x = "Candidate", y = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggplot(cleaned_data, aes(x = State, y = Percentage)) +
  geom_boxplot() +
  labs(title = "Percentage by State", x = "State", y = "Percentage") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

numerical_data <- cleaned_data %>%
  select(PollScore, SampleSize, Percentage)

correlation_matrix <- cor(numerical_data, use = "complete.obs")

library(ggcorrplot)
ggcorrplot(correlation_matrix, lab = TRUE, title = "Correlation Matrix")
```

# Discussion

# Appendix

## Additional Data Details

## Model Details {#sec-model-details}

\newpage

# References

### 2024-10-22 scripts/00-simulate_data.R

@@ -1,54 +1,51 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Deyi Kong
-# Date: 17 October 2024
-# Contact: deyi.kong@mail.utoronto.ca
+# Purpose: 
+# Author: Sophia Brothers
+# Date: 
+# Contact: 
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `election_folder` rproj
+# Any other information needed?
 
+library(dplyr)
 
-#### Workspace setup ####
-library(tidyverse)
-set.seed(777)
+set.seed(123)
+pollster_ids <- c(1001, 1002, 1003, 1004)
+pollsters <- c("Pollster A", "Pollster B", "Pollster C", "Pollster D")
+sponsors <- c("Sponsor 1", "Sponsor 2", "Sponsor 3")
+states <- c("Florida", "Ohio", "Texas", "Georgia", "Arizona", "California")
+methodologies <- c("Phone", "Online", "Mixed")
+candidates <- c("Candidate 1", "Candidate 2", "Candidate 3")
+candidate_parties <- c("Party A", "Party B", "Party C")
+population_types <- c("Registered Voters", "Likely Voters", "Adults")
+n <- 100 
 
+start_dates <- sample(seq(as.Date('2024-01-01'), as.Date('2024-12-31'), by = "day"), n, replace = TRUE)
+end_dates <- start_dates + sample(1:30, n, replace = TRUE)
 
-#### Simulate data ####
-# Top 10 population States names
-states <- c(
-  "California",
-  "Texas",
-  "Florida",
-  "New York",
-  "Pennsylvania",
-  "Illinois",
-  "Ohio",
-  "Georgia",
-  "North Carolina",
-  "Michigan"
+poll_data <- data.frame(
+  poll_id = 1:n,
+  pollster_id = sample(pollster_ids, n, replace = TRUE),
+  pollster = sample(pollsters, n, replace = TRUE),
+  sponsor_ids = sample(1:3, n, replace = TRUE),
+  sponsors = sample(sponsors, n, replace = TRUE),
+  pollster_rating_id = sample(1:5, n, replace = TRUE),
+  numeric_grade = round(runif(n, 2, 5), 2),
+  pollscore = round(runif(n, 60, 100), 2),
+  methodology = sample(methodologies, n, replace = TRUE),
[…]

### 2024-10-22 scripts/01-test_simulated_data.R

@@ -1,97 +1,98 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Deyi Kong
-# Date: 17 October 2024
-# Contact: deyi.kong@mail.utoronto.ca
+# Purpose: 
+# Author: Sophia Brothers
+# Date: 
+# Contact: 
 # License: MIT
-# Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `election_folder` rproj
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed?
 
+library(dplyr)
+library(purrr)
 
-#### Workspace setup ####
-library(tidyverse)
+data <- read_csv(here::here("data/simulated_data/simulated_data.csv"))
 
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+test_results <- list()
 
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
+test_structure <- function(data) {
+  if (!is.data.frame(data)) {
+    return("poll_data is not a data frame.")
+  }
+  if (ncol(data) != 22) {
+    return("poll_data does not have the correct number of columns.")
+  }
+  if (nrow(data) != 100) {
+    return("poll_data does not have the correct number of rows.")
+  }
+  return("Structure test passed.")
 }
 
-
-#### Test data ####
-
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
-} else {
-  stop("Test Failed: The dataset does not have 151 rows.")
+test_column_names <- function(data) {
+  expected_colnames <- c("poll_id", "pollster_id", "pollster", "sponsor_ids", 
+                         "sponsors", "pollster_rating_id", "numeric_grade", 
+                         "pollscore", "methodology", "state", 
+                         "start_date", "end_date", "sponsor_candidate", 
+                         "sponsor_candidate_party", "endorsed_candidate_name", 
+                         "endorsed_candidate_party", "sample_size", 
+                         "population", "election_date", "stage", 
+                         "candidate_name", "pct")
+  if (!all(colnames(data) == expected_colnames)) {
+    return("Column names do not match expected names.")
+  }
[…]

### 2024-10-22 scripts/02-download_data.R

@@ -1,20 +1,13 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from https://projects.fivethirtyeight.com/polls/president-general/2024/national/
-# Author: Deyi Kong
-# Date: 17 October 2024
-# Contact: deyi.kong@mail.utoronto.ca
+# Purpose: 
+# Author: Sophia Brothers
+# Date: 
+# Contact: 
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed and loaded
-# Any other information needed? Make sure you are in the `election_folder` rproj
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed?
 
 
-#### Workspace setup ####
-library(tidyverse)
+library(readr)
 
-#### Download data ####
-data <- read.csv("/Users/a._./Downloads/president_polls.csv")
-
-#### Save data ####
-write_csv(data, "data/01-raw_data/raw_data.csv") 
-
-         
+data <- read_csv(here::here("data/raw_data/president_polls.csv"))

### 2024-10-22 scripts/03-clean_data.R

@@ -1,0 +1,80 @@
+#### Preamble ####
+# Purpose: 
+# Author: Sophia Brothers
+# Date: 
+# Contact: 
+# License: MIT
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed?
+ 
+library(tidyverse)
+library(here)
+
+data <- read_csv(here::here("data/raw_data/president_polls.csv"))
+print(head(data))
+cleaned_data <- tryCatch({
+  data %>%
+    select(
+      poll_id,
+      pollster_id,
+      pollster,
+      party,
+      sponsors,
+      numeric_grade,
+      pollscore,
+      methodology,
+      state,
+      start_date,
+      end_date,
+      sample_size,
+      population,
+      election_date,
+      stage,
+      candidate_name,
+      pct
+    ) %>%
+  
+    rename(
+      Pollster = pollster,
+      Party = party,
+      NumericGrade = numeric_grade,
+      PollScore = pollscore,
+      Methodology = methodology,
+      State = state,
+      StartDate = start_date,
+      EndDate = end_date,
+      SampleSize = sample_size,
+      Population = population,
+      ElectionDate = election_date,
+      Stage = stage,
+      CandidateName = candidate_name,
+      Percentage = pct
+    ) %>%
+  
+    mutate(
+      # Specify the format of the date for proper parsing
+      StartDate = mdy(StartDate),  # Use mdy() for MM/DD/YYYY format
+      EndDate = mdy(EndDate),      # Same here
+      ElectionDate = mdy(ElectionDate)  # And here
+    ) %>%
+    
+    filter(!is.na(StartDate), !is.na(EndDate), !is.na(ElectionDate)) %>%
+    group_by(State, StartDate, Pollster) %>%
+    mutate(
+      CandidateName = ifelse(CandidateName == "Joe Biden", 
+                             "Kamala Harris", CandidateName)
+    ) %>%
+    ungroup()
+    }, 
+  error = function(e) {
+  message("An error occurred during data cleaning: ", e)
+  NULL
+})
+
+if (!is.null(cleaned_data)) {
+  write_csv(cleaned_data, here::here("data/analysis_data/president_polls_cleaned.csv"))
+  
+  print(head(cleaned_data))
+} else {
+  message("Cleaned data not available for saving due to an error.")
+}

### 2024-10-22 scripts/04-test_analysis_data.R

@@ -1,0 +1,97 @@
+#### Preamble ####
+# Purpose: 
+# Author: Sophia Brothers
+# Date: 
+# Contact: 
+# License: MIT
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed?
+
+library(tidyverse)
+library(here)
+
+cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv"))
+
+tests <- list(
+  structure_test = function(data) {
+    if (!inherits(data, "tbl_df")) {
+      stop("Structure test failed: Data is not a tibble.")
+    }
+    if (ncol(data) == 0) {
+      stop("Structure test failed: Data has no columns.")
+    }
+    if (nrow(data) == 0) {
+      stop("Structure test failed: Data has no rows.")
+    }
+    cat("Structure test passed.\n")
+  },
+  
+  column_names_test = function(data) {
+    expected_colnames <- c("poll_id", "pollster_id", "Pollster", "Party", "sponsors", 
+                           "NumericGrade", "PollScore", "Methodology", 
+                           "State", "StartDate", "EndDate", 
+                           "SampleSize", "Population", "ElectionDate", 
+                           "Stage", "CandidateName", "Percentage")
+    
+    if (!all(names(data) == expected_colnames)) {
+      stop("Column names test failed: Column names do not match expected names.")
+    }
+    cat("Column names test passed.\n")
+  },
+  
+  na_values_test = function(data) {
+    if (any(is.na(data$pollscore))) {
+      stop("NA values test failed: pollscore contains NA values.")
+    }
+    if (any(is.na(data$candidate_name))) {
+      stop("NA values test failed: candidate_name contains NA values.")
+    }
+    cat("NA values test passed.\n")
+  },
+  
+  numeric_grade_pollscore_test = function(data) {
+    if (any(data$numeric_grade < 2 | data$numeric_grade > 5)) {
+      stop("Numeric grade test failed: numeric_grade is out of expected range.")
+    }
+    if (any(data$pollscore < 60 | data$pollscore > 100)) {
+      stop("Pollscore test failed: pollscore is out of expected range.")
+    }
+    cat("Numeric grade and pollscore test passed.\n")
+  },
+  
+  sample_size_test = function(data) {
+    if (any(data$sample_size < 500 | data$sample_size > 2000)) {
+      stop("Sample size test failed: sample_size is out of expected range.")
+    }
[…]

### 2024-10-22 scripts/05-exploratory_data_analysis.R

@@ -1,0 +1,31 @@
+#### Preamble ####
+# Purpose: 
+# Author: Sophia Brothers
+# Date: 
+# Contact: 
+# License: MIT
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed? Make sure you are in the `starter_folder` rproj
+
+library(tidyverse)
+library(janitor)
+library(here)
+library(reshape2)
+
+cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv")) |> 
+  clean_names()
+
+glimpse(cleaned_data)
+
+head(cleaned_data)
+
+tail(cleaned_data)
+
+cleaned_data |> 
+  slice_sample(n = 6)
+
+summary(cleaned_data)
+
+missing_data <- cleaned_data |> 
+  summarise(across(everything(), ~sum(is.na(.))))
+missing_data

### 2024-10-22 scripts/06-model_data.R

@@ -1,0 +1,31 @@
+library(tidyverse)
+library(here)
+
+cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv"))
+
+cleaned_data <- cleaned_data %>%
+  mutate(
+    CandidateName = as.factor(CandidateName),
+    State = as.factor(State),
+    Pollster = as.factor(Pollster),
+    StartDate = as.Date(StartDate),
+    EndDate = as.Date(EndDate),
+    Party = as.factor(Party),
+    Methodology = as.factor(Methodology),
+    NumericGrade = as.numeric(NumericGrade),
+    SampleSize = as.numeric(SampleSize)
+  )
+
+cleaned_data <- cleaned_data %>%
+  mutate(PollRecency = as.numeric(Sys.Date() - StartDate))
+
+cleaned_data <- cleaned_data %>%
+ filter(CandidateName %in% c("Donald Trump", "Kamala Harris"))
+
+lm_model <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + State + PollRecency + CandidateName, 
+               data = cleaned_data)
+
+summary(lm_model)
+
+write_csv(cleaned_data, here::here("data/analysis_data/model_data.csv"))
+saveRDS(lm_model, here::here("models/lm_model.rds"))

### 2024-10-22 scripts/07-electoral_data.R

@@ -1,0 +1,33 @@
+library(tidyverse)
+library(here)
+
+cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv"))
+cleaned_data <- cleaned_data %>%
+  mutate(
+    CandidateName = as.factor(CandidateName),
+    State = as.factor(State),
+    Pollster = as.factor(Pollster),
+    StartDate = as.Date(StartDate),
+    EndDate = as.Date(EndDate),
+    Party = as.factor(Party),
+    Methodology = as.factor(Methodology),
+    NumericGrade = as.numeric(NumericGrade),
+    SampleSize = as.numeric(SampleSize)
+  )
+
+cleaned_data <- cleaned_data %>%
+  mutate(PollRecency = as.numeric(Sys.Date() - StartDate))
+
+cleaned_data <- cleaned_data %>%
+  filter(CandidateName %in% c("Donald Trump", "Kamala Harris")) %>%
+  select(State, PollScore, SampleSize, NumericGrade, PollRecency, Percentage, CandidateName)
+
+lm_model_trump <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State, 
+                     data = cleaned_data %>% filter(CandidateName == "Donald Trump"))
+
+lm_model_harris <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State, 
+                      data = cleaned_data %>% filter(CandidateName == "Kamala Harris"))
+
+write_csv(cleaned_data, here::here("data/analysis_data/electoral_data.csv"))
+saveRDS(lm_model_trump, here::here("models/lm_model_trump.rds"))
+saveRDS(lm_model_harris, here::here("models/lm_model_harris.rds"))

### 2024-10-22 paper/paper.qmd
Add files via upload

@@
 ---
-title: "Modeling the 2024 US Presidential Elections"
-subtitle: "My subtitle if needed"
+title: "Trump vs. Harris: A Data-Driven Forecast modeling the 2024 US Presidential Elections"
 author: 
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This study models the 2024 U.S.
+Presidential election outcomes using advanced statistical techniques and predictive modeling.
+The analysis incorporates data from various polling sources and considers key demographic and political factors to forecast both the popular vote and Electoral College results.
+The findings suggest a close race between the two main candidates, Donald Trump and Kamala Harris, with a predicted victory for Trump in the Electoral College.
+This analysis contributes to understanding how public opinion data and electoral systems interact, providing insights into electoral dynamics in contemporary U.S.
+politics."
 format: pdf
@@
 
-Quarto enables you to weave together content and executable code into a finished document. To learn more about Quarto see <https://quarto.org>.
+Elections in the United States, particularly presidential elections, are pivotal moments that shape not only the nation’s political landscape but also its international relations.
+The 2024 U.S.
+Presidential election, like its predecessors, promises to be contentious and closely watched.
+As electoral outcomes have become increasingly difficult to predict due to polarized political climates and fluctuating public opinion, sophisticated statistical models are essential tools for forecasting results and understanding the dynamics at play.
 
-# Data
+The core estimand of this study is the projected electoral outcome in terms of state-level victories and Electoral College votes for each candidate.
+The model leverages polling averages, historical voting patterns, and demographic predictors to estimate the probability of each candidate winning in individual states.
+The objective is to translate these probabilities into a forecast of Electoral College votes, thus determining the likely overall winner of the election.
+Given the significant influence of swing states, the model focuses on capturing their dynamics more accurately than national polling averages typically allow.
+
+The results suggest a close race, with Trump holding a slight edge in the Electoral College due to projected victories in key swing states, despite Harris maintaining a lead in national popular vote estimates.
+Trump's path to victory hinges on critical wins in states such as Pennsylvania, Georgia, and Wisconsin, where his projected margins are narrow but statistically significant.
+Meanwhile, Harris secures strong performances in populous Democratic strongholds like California and New York, but these do not offset Trump’s advantage in less populous yet crucial battleground states.
+
+Understanding these projected outcomes is essential for several reasons.
+First, the model highlights the limitations of relying solely on national polling to predict election results, emphasizing the continued relevance of the Electoral College system in determining the presidency.
+Additionally, it sheds light on the evolving political landscape in swing states, where demographic changes and shifting voter allegiances can have outsized impacts.
+These insights contribute to a broader understanding of how statistical modeling can inform political strategy and election analysis.
+
+The structure of the paper is as follows: In @sec-data, we provide outlines the data sources and variables considered, followed by the model setup in [@sec-mod] and justification.
+The results [@sec-result] presents the key findings of the analysis, with a discussion on the implications of these results for political forecasting and public opinion analysis.
+Finally, the paper concludes with an overview of potential limitations and suggestions for future research in the discussion [@sec-discussion].
+
+# Data {#sec-data}
 
@@
 
-This data was analyzed using [@citeR]. We also used the packages: tidyverse [@tidyverse], etc
+The data used in this analysis comes from a combination of publicly available polling data for the 2024 U.S.
+Presidential election.
+The analysis leverages R and several libraries, including tidyverse [@tidyverse], janitor [@janitor], knitr [@knitr], dplyr [@dplyr], purrr [@purrr], and here [@here], for data manipulation, and ggplot2 [@ggplot] for visualization.
+The dataset covers various polls conducted across multiple states, capturing the support for each major candidate—Donald Trump and Kamala Harris—along with detailed attributes of the polls.
 
 ## Measurement
+The primary measurements in the dataset include polling scores, sample sizes, and predicted percentages for each candidate.
+The data is filtered to include only polls relevant to the two primary candidates, and features like PollRecency were constructed to weigh recent polls higher.
+The data was cleaned by removing irrelevant or incomplete observations, and new variables were constructed where necessary.
 
 ## Outcome Variables
+The main outcome variable of interest is the Percentage of support each candidate has in a given poll, which represents the proportion of respondents who favor one candidate over the other.
+This variable is crucial for estimating both the popular vote and Electoral College outcomes.
 
 ## Predictor Variables
+Key predictors include:
 
-# Model
+PollScore: A numeric value that reflects the credibility of the pollster.
+
+SampleSize: The number of respondents in each poll.
+
+State: The U.S. state where the poll was conducted.
+
+PollRecency: The number of days between the start date of the poll and the present, which gives more weight to recent polls.
+
+NumericGrade: A numerical representation of the reliability of the pollster.
+
+CandidateName: The name of the candidate being polled (either Donald Trump or Kamala Harris).
+
+# Model {#sec-mod}
+To predict the outcome of the 2024 U.S.
+Presidential election, we developed two linear regression models: one for Donald Trump and one for Kamala Harris.
+These models aim to estimate the percentage of support for each candidate based on the aforementioned predictor variables.
 
 ## Model Set-Up
+This paper utilized a linear model:
+$$Percentage_i = \beta_0 + \beta_1 Pollscore_i + \beta_2 SampleSize + \beta_3 Numeric Grade_i + \beta_4 State_i + \beta_5PollRecency + \epsilon_i$$
+where, Percentage$_i$ represents the predicted support percentage for a candidate in a given poll.
+
+$\beta_i$ are the coefficients of the model that quantify the impact of each predictor.
+
+$\epsilon_i$ is the error term.
+
+The results from these models are aggregated to compute the predicted percentage of support for each candidate in each state.
+Based on these percentages, we forecast the likely winner of the state’s electoral votes.
 
 ## Model Justification
+This model was chosen for its balance between simplicity and predictive power.
+Linear regression allows us to capture relationships between polling quality, sample size, and recency, while still providing interpretable coefficients.
+Given the nature of polling data, this approach is both appropriate and commonly used in election forecasting.
+The model was validated using out-of-sample testing and Root Mean Square Error to ensure predictive accuracy.
 
@@
 print(state_winners, n=54)
-
 ```
 
-# Results
+# Results {#sec-result}
+The model results (see Appendix for details) show a clear distribution of support for both candidates across various states.
+The predicted percentages of support were normalized to ensure they summed up to 100%.
+After computing these results, we forecast that Donald Trump is likely to win 302 electoral votes, while Kamala Harris is expected to secure 233 electoral votes.
 
+[WILL CHANGE THE FORMAT OF THESE GRAPHS AND CAPTION/DESCRIPTION WILL BE ADDED]
 ```{r}
+#| label: fig-sum
+#| fig-cap: [Need Edit]
+#| fig-subcap: [""]
 #| echo: false
@@
 
-# Discussion
+\newpage
+
+# Discussion {#sec-discussion}
+In this paper, we developed and implemented a predictive model to forecast the outcome of the 2024 U.S.
+Presidential election, focusing on both popular vote percentages and Electoral College results.
+The model incorporates polling data, demographic information, and state-specific effects.
+
+The model teaches us that state-level dynamics play a crucial role in determining the winner of U.S.
+Presidential elections.
+Even though Harris is expected to lead in the popular vote, Trump’s success in key battleground states suggests a victory in the Electoral College.
+An additional insight is that poll recency and pollster reliability significantly affect the accuracy of electoral predictions.
+By weighing recent and higher-quality polls more heavily, the model improves the robustness of its forecasts.
+
+While the model adopts a linear regression framework for its simplicity and interpretability, this choice may limit its ability to capture more complex relationships among variables.
+Future iterations might benefit from exploring nonlinear models or advanced machine learning techniques, which could uncover patterns not readily observable through traditional methods.
+
+One limitation of this model is that it depends entirely on polling data, which may not always capture late shifts in voter preferences or turnout variations.
+Furthermore, the model does not account for external factors like campaign events or last-minute developments, which can significantly influence electoral outcomes.
+
+Future research should aim to integrate alternative data sources, such as social media sentiment or economic indicators, to provide a more comprehensive view of voter preferences.
+More advanced machine learning techniques could also be explored to improve the accuracy of future election predictions.
+
+\newpage
 

### 2024-10-22 paper/paper.qmd
Updated paper to add appendix outline

@@
 ## Measurement
+
 The primary measurements in the dataset include polling scores, sample sizes, and predicted percentages for each candidate.
@@
 ## Outcome Variables
+
 The main outcome variable of interest is the Percentage of support each candidate has in a given poll, which represents the proportion of respondents who favor one candidate over the other.
@@
 ## Predictor Variables
+
 Key predictors include:
@@
 # Model {#sec-mod}
+
 To predict the outcome of the 2024 U.S.
@@
 ## Model Set-Up
-This paper utilized a linear model:
-$$Percentage_i = \beta_0 + \beta_1 Pollscore_i + \beta_2 SampleSize + \beta_3 Numeric Grade_i + \beta_4 State_i + \beta_5PollRecency + \epsilon_i$$
-where, Percentage$_i$ represents the predicted support percentage for a candidate in a given poll.
+
+This paper utilized a linear model: $$Percentage_i = \beta_0 + \beta_1 Pollscore_i + \beta_2 SampleSize + \beta_3 Numeric Grade_i + \beta_4 State_i + \beta_5PollRecency + \epsilon_i$$ where, Percentage$_i$ represents the predicted support percentage for a candidate in a given poll.
 
@@
 ## Model Justification
+
 This model was chosen for its balance between simplicity and predictive power.
@@
 # Results {#sec-result}
+
 The model results (see Appendix for details) show a clear distribution of support for both candidates across various states.
@@
 
-[WILL CHANGE THE FORMAT OF THESE GRAPHS AND CAPTION/DESCRIPTION WILL BE ADDED]
+\[WILL CHANGE THE FORMAT OF THESE GRAPHS AND CAPTION/DESCRIPTION WILL BE ADDED\]
+
 ```{r}
@@
 print(summary_stats)
+```
 
+```{r}
+#| label: fig-pollscore
+#| fig-cap: [Need Edit]
+#| fig-subcap: [""]
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 ggplot(cleaned_data, aes(x = PollScore)) +
@@
   labs(title = "Distribution of Poll Scores", x = "Poll Score", y = "Frequency")
+```
 
+```{r}
+#| label: fig-samplesize
+#| fig-cap: [Need Edit]
+#| fig-subcap: [""]
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 ggplot(cleaned_data, aes(x = SampleSize)) +
@@
   labs(title = "Distribution of Sample Size", x = "Sample Size", y = "Frequency")
+```
 
+```{r}
+#| label: fig-percentage
+#| fig-cap: [Need Edit]
+#| fig-subcap: [""]
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 ggplot(cleaned_data, aes(x = Percentage)) +
@@
   labs(title = "Distribution of Percentage", x = "Percentage", y = "Frequency")
+```
 
+```{r}
+#| label: fig-candidatepercentage
+#| fig-cap: [Need Edit]
+#| fig-subcap: [""]
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 ggplot(cleaned_data, aes(x = CandidateName, y = Percentage)) +
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
 
+will update these box plots to different visuals due to box plots not showing the full dataset!
+
+```{r}
+#| label: fig-statepercentage
+#| fig-cap: [Need Edit]
+#| fig-subcap: [""]
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 ggplot(cleaned_data, aes(x = State, y = Percentage)) +
@@
   theme(axis.text.x = element_text(angle = 45, hjust = 1))
+```
 
+```{r}
+#| label: fig-corrmatrix
+#| fig-cap: [Need Edit]
+#| fig-subcap: [""]
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 numerical_data <- cleaned_data %>%
@@
 # Discussion {#sec-discussion}
+
 In this paper, we developed and implemented a predictive model to forecast the outcome of the 2024 U.S.
@@
 
+## Idealized Methodology
+
+This is an outline. I will turn this into paragraph style this week.
+
+**Target Population:** The target population is eligible US voters from all 50 states and Washington D.C. –\> representation from every state for an Electoral College forecast.
+
+**Sampling Method**
+
+\- Stratified Random Sampling: The sample will be stratified by state, ensuring each state is proportionally represented in terms of population size and key voter demographics (age, gender, race/ethnicity, income, political affiliation).
+
+\- Key Stratification Variables: Within each state, voters will be further stratified based on party affiliation, urban/rural distribution, and voting history.
+
+\- For swing states, we will ensure oversampling to capture more granular data.
+
+**Sample Size**
+
+\- The total sample size will be 15,000 respondents –\> ensures state-level representation while maintaining a national margin of error of ±1%.
+
+\- As mentioned above, swing states will have an additional focus, with 1,000+ respondents per state (depending on population size).
+
+\- Electoral vote-rich states (e.g., California, Texas, Florida) will be proportionally represented to reflect their impact on the election outcome.
+
+\- With an estimated response rate of 10%, we aim to recruit 150,000 potential respondents across states.
+
+**Recruiting Respondents**
+
+\- Panel Providers: Utilize online panel providers (such as YouGov, Ipsos, or Dynata) for geographically and demographically targeted respondents by state.
+
+\- Social Media Ads: Conduct state-targeted ads on platforms like Facebook and Instagram\*for broad geographic reach
+
+\- State-Specific Phone Surveys: To reach older and less tech-savvy voters, particularly in states with higher proportions of rural populations, we will use IVR phone surveys and live calls.
+
+\- Respondents will be entered into state-level sweepstakes, with one winner per state
+
+**Data Collection**
+
+\- Surveys will be conducted on platforms like Qualtrics or SurveyMonkey, with branching logic for state-specific questions.
+Telephone surveys for older demographics will be integrated.
+The mock survey will be conducted using Google Forms for the purposes of accessibility.
+
+\- The survey will take approximately 5 minutes in order to minimize drop-off rates
+
+\- The survey will include state-specific questions and national-level questions, such as voting intent, candidate preference, key issues, and voter perception.
+
+\- Weighting by State: Data will be weighted post-collection to ensure state-level accuracy, considering the latest Census data and state voting records.
+
+**Data Validation**
+
+\- Validate completion time (speeding checks), IP tracking (to prevent duplicate responses), and reCAPTCHA to prevent bots.
+
+\- Respondents will be asked their state of residence, and ZIP code for additional verification.
+
+\- Data will be weighted by state demographics and previous election turnout using Census and state voting data.
+Swing states will receive additional attention to ensure no underrepresentation of key voter segments.
+
+include budget allocation?
+
+**Survey Structure (will be migrated to google form)**
+
+Introduction
+
+\- Brief description of the survey’s purpose (US Presidential Election Forecast with a focus on state-level dynamics).
+
+\- Assurance of confidentiality and anonymity, with a clear outline of data use.
+
+\- Contact details for survey-related inquiries.
+
+Questions
+
+1\. Demographics:
+
+\- State of Residence (Dropdown list of all 50 states and D.C.)
+
+\- Age, gender, race/ethnicity, income level, education, party affiliation (Republican, Democrat, Independent, Other).
+
+2\. Voter Registration:
+
+\- Are you registered to vote in the upcoming election? (Yes/No)
+
+\- How likely are you to vote in the upcoming election? (Likert scale)
+
+3\. Voting Intent:
+
+\- If the election were held today, who would you vote for? (List of candidates specific to the respondent’s state, based on ballot access).
+
+\- How certain are you of your choice? (Likert scale)
+
+4\. Key Issues:
+
+\- What are the most important issues influencing your vote? (Rank the top 3 issues: Economy, Healthcare, Immigration, Climate Change, National Security, Social Issues, etc.)
+
+5\. State-Specific Issues:
+
+\- What state-level issues are important to you? (Open-ended question, with suggestions like state economy, healthcare access, local infrastructure, etc.)\
+
+6\. Perceptions of Candidates:
+
+\- Rate the following candidates on trustworthiness, competence, leadership, etc. (Likert scale, customized for state-level candidates where applicable).
+
+7\. Previous Voting Behavior:
+
+\- Did you vote in the last presidential election? (Yes/No)
+
+\- If yes, who did you vote for? (Open response, with drop-down of major candidates for validation)
+
+Closing
+
+\- Thank You: Thank the respondents for their time and participation, reminding them of the sweepstakes.
+
 \newpage

### 2024-10-22 scripts/00-simulate_data.R
Updated scripts to add comments

@@ -1,15 +1,19 @@
 #### Preamble ####
-# Purpose: 
+# Purpose: Simulate polling data for the 2024 US Election for use in election forecasting
 # Author: Sophia Brothers
-# Date: 
-# Contact: 
+# Date: October 22nd, 2024
+# Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed?
+# Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
+# load libraries
 library(dplyr)
 
+# set seed for reproducability
 set.seed(123)
+
+# define possible values to simulate polling data
 pollster_ids <- c(1001, 1002, 1003, 1004)
 pollsters <- c("Pollster A", "Pollster B", "Pollster C", "Pollster D")
 sponsors <- c("Sponsor 1", "Sponsor 2", "Sponsor 3")
@@ -18,11 +22,13 @@ methodologies <- c("Phone", "Online", "Mixed")
 candidates <- c("Candidate 1", "Candidate 2", "Candidate 3")
 candidate_parties <- c("Party A", "Party B", "Party C")
 population_types <- c("Registered Voters", "Likely Voters", "Adults")
-n <- 100 
+n <- 100
 
-start_dates <- sample(seq(as.Date('2024-01-01'), as.Date('2024-12-31'), by = "day"), n, replace = TRUE)
+# generate random start and end dates for each poll
+start_dates <- sample(seq(as.Date("2024-01-01"), as.Date("2024-12-31"), by = "day"), n, replace = TRUE)
 end_dates <- start_dates + sample(1:30, n, replace = TRUE)
 
+# create a simulated data frame of polls with various attributes
 poll_data <- data.frame(
   poll_id = 1:n,
   pollster_id = sample(pollster_ids, n, replace = TRUE),
@@ -42,10 +48,11 @@ poll_data <- data.frame(
   endorsed_candidate_party = sample(candidate_parties, n, replace = TRUE),
   sample_size = sample(500:2000, n, replace = TRUE),
   population = sample(population_types, n, replace = TRUE),
-  election_date = sample(seq(as.Date('2024-11-01'), as.Date('2024-11-08'), by="day"), n, replace = TRUE),
+  election_date = sample(seq(as.Date("2024-11-01"), as.Date("2024-11-08"), by = "day"), n, replace = TRUE),
[…]

### 2024-10-22 scripts/01-test_simulated_data.R

@@ -1,19 +1,24 @@
 #### Preamble ####
-# Purpose: 
+# Purpose: This script reads in the simulated polling data, performs a series of validation tests
+# to ensure the data structure, column names, and values are correct, and reports the results.
 # Author: Sophia Brothers
-# Date: 
-# Contact: 
+# Date: October 22nd, 2024
+# Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed?
+# Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
+# load libraries
 library(dplyr)
 library(purrr)
 
+# read the simulated data
 data <- read_csv(here::here("data/simulated_data/simulated_data.csv"))
 
+# initialize an empty list to store the results of each test (pass/fail)
 test_results <- list()
 
+# test if the data structure is correct
 test_structure <- function(data) {
   if (!is.data.frame(data)) {
     return("poll_data is not a data frame.")
@@ -27,21 +32,25 @@ test_structure <- function(data) {
   return("Structure test passed.")
 }
 
+# test if the column names match the expected column names
 test_column_names <- function(data) {
-  expected_colnames <- c("poll_id", "pollster_id", "pollster", "sponsor_ids", 
-                         "sponsors", "pollster_rating_id", "numeric_grade", 
-                         "pollscore", "methodology", "state", 
-                         "start_date", "end_date", "sponsor_candidate", 
-                         "sponsor_candidate_party", "endorsed_candidate_name", 
-                         "endorsed_candidate_party", "sample_size", 
-                         "population", "election_date", "stage", 
-                         "candidate_name", "pct")
+  expected_colnames <- c(
+    "poll_id", "pollster_id", "pollster", "sponsor_ids",
+    "sponsors", "pollster_rating_id", "numeric_grade",
+    "pollscore", "methodology", "state",
+    "start_date", "end_date", "sponsor_candidate",
+    "sponsor_candidate_party", "endorsed_candidate_name",
+    "endorsed_candidate_party", "sample_size",
+    "population", "election_date", "stage",
+    "candidate_name", "pct"
+  )
   if (!all(colnames(data) == expected_colnames)) {
     return("Column names do not match expected names.")
   }
   return("Column names test passed.")
 }
 
[…]

### 2024-10-22 scripts/02-download_data.R

@@ -1,13 +1,14 @@
 #### Preamble ####
-# Purpose: 
+# Purpose: This script loads the raw presidential polling data from a CSV file for further analysis.
 # Author: Sophia Brothers
-# Date: 
-# Contact: 
+# Date: October 22nd, 2024
+# Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed?
+# Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
-
+# load library
 library(readr)
 
+# Read the raw presidential polls data
 data <- read_csv(here::here("data/raw_data/president_polls.csv"))

### 2024-10-22 scripts/03-clean_data.R

@@ -1,79 +1,85 @@
 #### Preamble ####
-# Purpose: 
+# Purpose: This script reads and cleans raw presidential polling data,
+# performing various transformations, and outputs the cleaned data to a CSV file.
 # Author: Sophia Brothers
-# Date: 
-# Contact: 
+# Date: October 22nd, 2024
+# Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed?
- 
+# Pre-requisites: The `tidyverse` and 'here' package must be installed
+# Any other information needed? Make sure you are in the `US_Election_2024` rproj
+
+# load libraries
 library(tidyverse)
 library(here)
 
+# read data
 data <- read_csv(here::here("data/raw_data/president_polls.csv"))
-print(head(data))
-cleaned_data <- tryCatch({
-  data %>%
-    select(
-      poll_id,
-      pollster_id,
-      pollster,
-      party,
-      sponsors,
-      numeric_grade,
-      pollscore,
-      methodology,
-      state,
-      start_date,
-      end_date,
-      sample_size,
-      population,
-      election_date,
-      stage,
-      candidate_name,
-      pct
-    ) %>%
-  
-    rename(
-      Pollster = pollster,
-      Party = party,
-      NumericGrade = numeric_grade,
-      PollScore = pollscore,
-      Methodology = methodology,
-      State = state,
-      StartDate = start_date,
-      EndDate = end_date,
-      SampleSize = sample_size,
-      Population = population,
-      ElectionDate = election_date,
-      Stage = stage,
-      CandidateName = candidate_name,
-      Percentage = pct
-    ) %>%
-  
-    mutate(
-      # Specify the format of the date for proper parsing
-      StartDate = mdy(StartDate),  # Use mdy() for MM/DD/YYYY format
-      EndDate = mdy(EndDate),      # Same here
-      ElectionDate = mdy(ElectionDate)  # And here
-    ) %>%
-    
-    filter(!is.na(StartDate), !is.na(EndDate), !is.na(ElectionDate)) %>%
-    group_by(State, StartDate, Pollster) %>%
-    mutate(
-      CandidateName = ifelse(CandidateName == "Joe Biden", 
-                             "Kamala Harris", CandidateName)
-    ) %>%
-    ungroup()
-    }, 
-  error = function(e) {
-  message("An error occurred during data cleaning: ", e)
-  NULL
-})
 
[…]

### 2024-10-22 scripts/04-test_analysis_data.R

@@ -1,18 +1,22 @@
 #### Preamble ####
-# Purpose: 
+# Purpose: This script performs a series of validation tests on cleaned presidential polling data.
 # Author: Sophia Brothers
-# Date: 
-# Contact: 
+# Date: October 22nd, 2024
+# Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed?
+# Pre-requisites: The `tidyverse` and 'here' package must be installed
+# Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
+# load libraries
 library(tidyverse)
 library(here)
 
+# load data
 cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv"))
 
+# define list of tests
 tests <- list(
+  # test structure of data
   structure_test = function(data) {
     if (!inherits(data, "tbl_df")) {
       stop("Structure test failed: Data is not a tibble.")
@@ -25,72 +29,66 @@ tests <- list(
     }
     cat("Structure test passed.\n")
   },
-  
+  # test if column names match expected names
   column_names_test = function(data) {
-    expected_colnames <- c("poll_id", "pollster_id", "Pollster", "Party", "sponsors", 
-                           "NumericGrade", "PollScore", "Methodology", 
-                           "State", "StartDate", "EndDate", 
-                           "SampleSize", "Population", "ElectionDate", 
-                           "Stage", "CandidateName", "Percentage")
-    
+    expected_colnames <- c(
+      "poll_id", "pollster_id", "Pollster", "Party", "sponsors",
+      "NumericGrade", "PollScore", "Methodology",
+      "State", "StartDate", "EndDate",
+      "SampleSize", "Population", "ElectionDate",
+      "Stage", "CandidateName", "Percentage"
+    )
+
     if (!all(names(data) == expected_colnames)) {
       stop("Column names test failed: Column names do not match expected names.")
     }
     cat("Column names test passed.\n")
   },
-  
+  # test if certain columns contain NA values
   na_values_test = function(data) {
-    if (any(is.na(data$pollscore))) {
-      stop("NA values test failed: pollscore contains NA values.")
+    if (any(is.na(data$Percentage))) {
+      stop("NA values test failed: Percentage contains NA values.")
     }
-    if (any(is.na(data$candidate_name))) {
[…]

### 2024-10-22 scripts/05-exploratory_data_analysis.R

@@ -1,31 +1,40 @@
 #### Preamble ####
-# Purpose: 
+# Purpose: Perform exploratory data analysis on cleaned data
 # Author: Sophia Brothers
-# Date: 
-# Contact: 
+# Date:
+# Contact:
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Pre-requisites: The `tidyverse`, 'janitor', 'here', and 'reshape2' packages
+# must be installed
+# Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
+# load libraries
 library(tidyverse)
 library(janitor)
 library(here)
 library(reshape2)
 
-cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv")) |> 
+# read cleaned data
+cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv")) |>
   clean_names()
 
+# quick overview of data structure
 glimpse(cleaned_data)
 
+# print first few rows of data
 head(cleaned_data)
 
+# print last few rows of data
 tail(cleaned_data)
 
-cleaned_data |> 
+# randomly sample 6 rows
+cleaned_data |>
   slice_sample(n = 6)
 
+# summary of the dataset
 summary(cleaned_data)
 
-missing_data <- cleaned_data |> 
-  summarise(across(everything(), ~sum(is.na(.))))
+# number of NA values in each column
+missing_data <- cleaned_data |>
+  summarise(across(everything(), ~ sum(is.na(.))))
 missing_data

### 2024-10-22 scripts/06-model_data.R

@@ -1,8 +1,20 @@
+#### Preamble ####
+# Purpose: Run a linear regression model on president polls data
+# Author: Sophia Brothers
+# Date: October 22nd, 2024
+# Contact: sophia.brothers@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse` and 'here' packages must be installed
+# Any other information needed? Make sure you are in the `US_Election_2024` rproj
+
+# load libraries
 library(tidyverse)
 library(here)
 
+# load data
 cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv"))
 
+# convert some columns to the appropriate data types
 cleaned_data <- cleaned_data %>%
   mutate(
     CandidateName = as.factor(CandidateName),
@@ -16,16 +28,25 @@ cleaned_data <- cleaned_data %>%
     SampleSize = as.numeric(SampleSize)
   )
 
+# create a new variable called PollRecency
+# PollRecency measures how recently the poll was conducted (based on start date)
 cleaned_data <- cleaned_data %>%
   mutate(PollRecency = as.numeric(Sys.Date() - StartDate))
 
+# only consider polls for Trump or Harris
 cleaned_data <- cleaned_data %>%
- filter(CandidateName %in% c("Donald Trump", "Kamala Harris"))
+  filter(CandidateName %in% c("Donald Trump", "Kamala Harris"))
 
-lm_model <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + State + PollRecency + CandidateName, 
-               data = cleaned_data)
+# linear regression model that predicts support percentages
+lm_model <- lm(
+  Percentage ~ PollScore + SampleSize + NumericGrade + State +
+    PollRecency + CandidateName,
+  data = cleaned_data
+)
 
+# summary of model
 summary(lm_model)
 
+# save modified data for the model as a CSV and save model as RDS
 write_csv(cleaned_data, here::here("data/analysis_data/model_data.csv"))
 saveRDS(lm_model, here::here("models/lm_model.rds"))

### 2024-10-22 scripts/07-electoral_data.R

@@ -1,6 +1,17 @@
+#### Preamble ####
+# Purpose: Run a linear regression model to predict the electoral college
+# Author: Sophia Brothers
+# Date: October 22nd, 2024
+# Contact: sophia.brothers@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse` and 'here' packages must be installed
+# Any other information needed? Make sure you are in the `US_Election_2024` rproj
+
+# load libraries
 library(tidyverse)
 library(here)
 
+# convert some columns to the appropriate data types
 cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv"))
 cleaned_data <- cleaned_data %>%
   mutate(
@@ -15,19 +26,27 @@ cleaned_data <- cleaned_data %>%
     SampleSize = as.numeric(SampleSize)
   )
 
+# create a new variable called PollRecency
+# PollRecency measures how recently the poll was conducted (based on start date)
 cleaned_data <- cleaned_data %>%
   mutate(PollRecency = as.numeric(Sys.Date() - StartDate))
 
+# only consider polls for Trump or Harris
 cleaned_data <- cleaned_data %>%
   filter(CandidateName %in% c("Donald Trump", "Kamala Harris")) %>%
   select(State, PollScore, SampleSize, NumericGrade, PollRecency, Percentage, CandidateName)
 
-lm_model_trump <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State, 
-                     data = cleaned_data %>% filter(CandidateName == "Donald Trump"))
+# linear model for Donald Trump's poll percentage based on various predictors
+lm_model_trump <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State,
+  data = cleaned_data %>% filter(CandidateName == "Donald Trump")
+)
 
-lm_model_harris <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State, 
-                      data = cleaned_data %>% filter(CandidateName == "Kamala Harris"))
+# linear model for Kamala Harris' poll percentage based on various predictors
+lm_model_harris <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State,
+  data = cleaned_data %>% filter(CandidateName == "Kamala Harris")
+)
 
+# save modified data for the model as a CSV and save models as RDS
[…]

### 2024-10-22 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -4,11 +4,12 @@
 # Date: October 22nd, 2024
 # Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
+# Pre-requisites: The `tidyverse` and 'arrow' packages must be installed
 # Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
 # load libraries
 library(dplyr)
+library(arrow)
 
 # set seed for reproducability
 set.seed(123)
@@ -56,3 +57,4 @@ poll_data <- data.frame(
 
 # write the cleaned and simulated polling data to a CSV file
 write_csv(cleaned_data, here::here("data/simulated_data/simulated_data.csv"))
+write_parquet(cleaned_data, here::here("data/simulated_data/simulated_data.parquet"))

### 2024-10-22 scripts/01-test_simulated_data.R
Update 01-test_simulated_data.R

@@ -5,15 +5,16 @@
 # Date: October 22nd, 2024
 # Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
+# Pre-requisites: The `tidyverse` and 'arrow' packages must be installed
 # Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
 # load libraries
 library(dplyr)
 library(purrr)
+library(arrow)
 
 # read the simulated data
-data <- read_csv(here::here("data/simulated_data/simulated_data.csv"))
+data <- read_parquet(here::here("data/simulated_data/simulated_data.parquet"))
 
 # initialize an empty list to store the results of each test (pass/fail)
 test_results <- list()

### 2024-10-22 scripts/03-clean_data.R
Update 03-clean_data.R

@@ -5,12 +5,13 @@
 # Date: October 22nd, 2024
 # Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` and 'here' package must be installed
+# Pre-requisites: The `tidyverse`, 'here', and 'arrow' packages must be installed
 # Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
 # load libraries
 library(tidyverse)
 library(here)
+library(arrow)
 
 # read data
 data <- read_csv(here::here("data/raw_data/president_polls.csv"))
@@ -79,7 +80,8 @@ cleaned_data <- tryCatch(
 # save cleaned data as new csv if successful, print error if not
 if (!is.null(cleaned_data)) {
   write_csv(cleaned_data, here::here("data/analysis_data/president_polls_cleaned.csv"))
-
+  write_parquet(cleaned_data, here::here("data/analysis_data/president_polls_cleaned.parquet"))
+  
   print(head(cleaned_data))
 } else {
   message("Cleaned data not available for saving due to an error.")

### 2024-10-22 scripts/04-test_analysis_data.R
Update 04-test_analysis_data.R

@@ -4,15 +4,16 @@
 # Date: October 22nd, 2024
 # Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` and 'here' package must be installed
+# Pre-requisites: The `tidyverse`, 'here', and 'arrow' packages must be installed
 # Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
 # load libraries
 library(tidyverse)
 library(here)
+library(arrow)
 
 # load data
-cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv"))
+cleaned_data <- read_parquet(here::here("data/analysis_data/president_polls_cleaned.parquet"))
 
 # define list of tests
 tests <- list(

### 2024-10-22 scripts/05-exploratory_data_analysis.R
Update 05-exploratory_data_analysis.R

@@ -4,7 +4,7 @@
 # Date:
 # Contact:
 # License: MIT
-# Pre-requisites: The `tidyverse`, 'janitor', 'here', and 'reshape2' packages
+# Pre-requisites: The `tidyverse`, 'janitor', 'here', 'arrow', and 'reshape2' packages
 # must be installed
 # Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
@@ -13,9 +13,10 @@ library(tidyverse)
 library(janitor)
 library(here)
 library(reshape2)
+library(arrow)
 
 # read cleaned data
-cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv")) |>
+cleaned_data <- read_parquet(here::here("data/analysis_data/president_polls_cleaned.parquet")) |>
   clean_names()
 
 # quick overview of data structure

### 2024-10-22 scripts/06-model_data.R
Update 06-model_data.R

@@ -4,15 +4,16 @@
 # Date: October 22nd, 2024
 # Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` and 'here' packages must be installed
+# Pre-requisites: The `tidyverse`, 'here', and 'arrow' packages must be installed
 # Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
 # load libraries
 library(tidyverse)
 library(here)
+library(arrow)
 
 # load data
-cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv"))
+cleaned_data <- read_parquet(here::here("data/analysis_data/president_polls_cleaned.parquet"))
 
 # convert some columns to the appropriate data types
 cleaned_data <- cleaned_data %>%
@@ -47,6 +48,7 @@ lm_model <- lm(
 # summary of model
 summary(lm_model)
 
-# save modified data for the model as a CSV and save model as RDS
+# save modified data for the model as a CSV + Parquet and save model as RDS
 write_csv(cleaned_data, here::here("data/analysis_data/model_data.csv"))
+write_parquet(cleaned_data, here::here("data/analysis_data/model_data.parquet"))
 saveRDS(lm_model, here::here("models/lm_model.rds"))

### 2024-10-22 scripts/07-electoral_data.R
Update 07-electoral_data.R

@@ -4,15 +4,16 @@
 # Date: October 22nd, 2024
 # Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` and 'here' packages must be installed
+# Pre-requisites: The `tidyverse`, 'here', and 'arrow' packages must be installed
 # Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
 # load libraries
 library(tidyverse)
 library(here)
+library(arrow)
 
 # convert some columns to the appropriate data types
-cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv"))
+cleaned_data <- read_parquet(here::here("data/analysis_data/president_polls_cleaned.parquet"))
 cleaned_data <- cleaned_data %>%
   mutate(
     CandidateName = as.factor(CandidateName),
@@ -46,7 +47,8 @@ lm_model_harris <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollR
   data = cleaned_data %>% filter(CandidateName == "Kamala Harris")
 )
 
-# save modified data for the model as a CSV and save models as RDS
+# save modified data for the model as a CSV + Parquet and save models as RDS
 write_csv(cleaned_data, here::here("data/analysis_data/electoral_data.csv"))
+write_parquet(cleaned_data, here::here("data/analysis_data/electoral_data.parquet"))
 saveRDS(lm_model_trump, here::here("models/lm_model_trump.rds"))
 saveRDS(lm_model_harris, here::here("models/lm_model_harris.rds"))

### 2024-10-22 paper/paper.qmd
Update paper.qmd

@@
 library(here)
+library(arrow)
 
-cleaned_data <- read_csv(here::here("data/analysis_data/model_data.csv"))
+cleaned_data <- read_parquet(here::here("data/analysis_data/model_data.parquet"))
 lm_model <- readRDS(here::here("models/lm_model.rds"))
@@
 library(here)
+library(arrow)
 
-cleaned_data <- read_csv(here::here("data/analysis_data/electoral_data.csv"))
+cleaned_data <- read_parquet(here::here("data/analysis_data/electoral_data.parquet"))
 lm_model_trump <- readRDS(here::here("models/lm_model_trump.rds"))
@@
 #| message: false
-cleaned_data <- read_csv(here::here("data/analysis_data/president_polls_cleaned.csv"))
+cleaned_data <- read_parquet(here::here("data/analysis_data/president_polls_cleaned.parquet"))
 

### 2024-10-22 paper/paper.qmd
Updated paper

@@
   - Deyi Kong
+  - Rayan Awad Alim
 thanks: "Code and data are available at: [https://github.com/eeeee-cmd/US_Election/](https://github.com/eeeee-cmd/US_Election/)."

### 2024-10-28 paper/paper.qmd
Update Visualizations and clean up packages enviroment set up

--- 2024-10-22 paper/paper.qmd
@@
 title: "Trump vs. Harris: A Data-Driven Forecast modeling the 2024 US Presidential Elections"
+subtitle: "A close race but Trump, trumps Harris"
 author: 
@@
 ---
+
+```{r}
+#| include: false
+#| warning: false
+#| message: false
+
+# Install Packages if not downloaded:
+required_packages <- c(
+  "tidyverse", "sf", "here", "arrow", "ggplot2", "janitor", "purrr", "knitr", "kableExtra"
+) 
+for (p in required_packages) { 
+  if (!require(p, character.only = TRUE)) {
+    install.packages(p, character.only = TRUE)
+  }
+}
+
+library(tidyverse)
+library(sf)
+library(dplyr)
+library(here)
+library(arrow)
+library(ggplot2)
+library(janitor)
+library(purrr)
+library(knitr)
+library(kableExtra)
+```
 
@@
 The data used in this analysis comes from a combination of publicly available polling data for the 2024 U.S.
-Presidential election.
-The analysis leverages R and several libraries, including tidyverse [@tidyverse], janitor [@janitor], knitr [@knitr], dplyr [@dplyr], purrr [@purrr], and here [@here], for data manipulation, and ggplot2 [@ggplot] for visualization.
+Presidential election[@FiveThirtyEightPolls].
+The analysis leverages the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `janitor` [@janitor], `knitr` [@knitr], `dplyr` [@dplyr], `purrr` [@purrr], and `here` [@here], for data manipulation, and `ggplot2` [@ggplot] for visualization.
 The dataset covers various polls conducted across multiple states, capturing the support for each major candidate—Donald Trump and Kamala Harris—along with detailed attributes of the polls.
@@
 #| message: false
-library(tidyverse)
-library(dplyr)
-library(here)
-library(arrow)
 
@@
 
-print(results)
+results %>%
+  kable(
+    caption = "Predicted Percentage Vote for Donald Trump and Kamala Harris",
+    col.names = c("Candidate Name", "Average Predicted Percentage", "Normalized Percentage"),
+    digits = 2,
+    booktabs = TRUE
+  ) %>%
+  kable_styling(latex_options = c("striped", "scale_down"))
 ```
@@
 #| message: false
-
-library(tidyverse)
-library(here)
-library(arrow)
 
@@
 
-print(electoral_totals)
+electoral_totals %>%
+  kable(
+    caption = "Predicted Number of Electoral Votes for Donald Trump and Kamala Harris",
+    col.names = c("Trumps Total Electoral Votes", "Harriss Total Electoral Votes"),
+    digits = 2,
+    booktabs = TRUE
+  ) %>%
+  kable_styling(latex_options = c("striped", "scale_down"))
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 
@@
   )
+#print(state_winners, n=54)
 
-print(state_winners, n=54)
+state_winners %>%
+  slice(1:53) %>%
+  kable(
+    caption = "Predicted Percentage of Electoral Votes Per State",
+    col.names = c("State", "Trump Predicted %", "Harris Predicted %", "Winner"),
+    digits = 2,
+    booktabs = TRUE
+  ) %>%
+  kable_styling(latex_options = c("striped", "scale_down"))
 ```
@@
 ```{r}
-#| label: fig-sum
+#| label: tbl-sum
 #| fig-cap: [Need Edit]
-#| fig-subcap: [""]
 #| echo: false
@@
 
-print(summary_stats)
+# print(summary_stats)
+
+summary_stats %>%
+  kable(
+    caption = "Summary Statistics of Model Results",
+    col.names = c("Average Poll Score", "Average Sample Size", "Average Percentage", "Total Polls"),
+    digits = 3,
+    booktabs= TRUE
+  ) %>%
+  kable_styling(latex_options = c("striped", "scale_down"))
 ```
@@
 ```{r}
-#| label: fig-pollscore
-#| fig-cap: [Need Edit]
-#| fig-subcap: [""]
+#| label: fig-predictedscoremap
+#| fig-cap: "Mapped Model-predicted outsom for 2024 elections by State"
 #| echo: false
@@
 #| message: false
-ggplot(cleaned_data, aes(x = PollScore)) +
-  geom_histogram(bins = 30, fill = "blue", alpha = 0.7) +
-  labs(title = "Distribution of Poll Scores", x = "Poll Score", y = "Frequency")
+
+states_map <- map_data("state")
+
+electoral_data <- predictions_with_votes
+electoral_data$State <- tolower(electoral_data$State) # matching state names
+
+merged_data <- inner_join(states_map, electoral_data, by = c("region" = "State"))
+
+map <- ggplot(merged_data, aes(x = long, y = lat, group = group)) +
+  geom_polygon(aes(fill = Winner), color = "white") +
+  scale_fill_manual(values = c("Donald Trump" = "red", "Kamala Harris" = "blue")) +
+  labs(title = "2024 US Presidential Election Predictions by State",
+       fill = "Predicted Winner") +
+  theme_void() +
+  theme(legend.position = "right")
+
+map
+```
+
+```{r}
+#| label: fig-pollscoredistribution
+#| fig-cap: Distribution of Poll Scores by Candidate with denisty overlayed
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+Select_candidates_cleaned_data <- cleaned_data %>%
+  filter(CandidateName %in% c("Donald Trump", "Kamala Harris"))
+
+ggplot(Select_candidates_cleaned_data, aes(x = PollScore)) +
+  geom_histogram(aes(y = ..density..),
+                 bins = 30,
+                 fill = "blue",
+                 color = "black",
+                 alpha = 0.7) +
+    geom_density(color = "black", fill = "salmon", alpha = 0.6) +
+  labs(title = "Distribution of Poll Scores by Candidate",
+       x = "Poll Score", 
+       y = "Frequency") +
+  theme_minimal() +
+  facet_wrap(~ CandidateName)
 ```
@@
 #| fig-cap: [Need Edit]
-#| fig-subcap: [""]
 #| echo: false
@@
 #| message: false
+
 ggplot(cleaned_data, aes(x = SampleSize)) +
-  geom_histogram(bins = 30, fill = "green", alpha = 0.7) +
-  labs(title = "Distribution of Sample Size", x = "Sample Size", y = "Frequency")
+  geom_histogram(bins = 30,
+                 fill = "green",
+                 color = "black",
+                 alpha = 0.7) +
+  labs(title = "Distribution of Sample Size",
+       x = "Sample Size", 
+       y = "Frequency")+
+  theme_minimal()
 ```
@@
 ggplot(cleaned_data, aes(x = Percentage)) +
-  geom_histogram(bins = 30, fill = "orange", alpha = 0.7) +
-  labs(title = "Distribution of Percentage", x = "Percentage", y = "Frequency")
+  geom_histogram(bins = 30,
+                 fill = "orange",
+                 color = "black",
+                 alpha = 0.7) +
+  labs(title = "Distribution of Percentage", 
+       x = "Percentage",
+       y = "Frequency") +
+  theme_minimal()
 ```
@@
 #| message: false
-ggplot(cleaned_data, aes(x = CandidateName, y = Percentage)) +
-  geom_boxplot() +
-  labs(title = "Percentage by Candidate", x = "Candidate", y = "Percentage") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+
+harris_trump_selection <- cleaned_data %>%
+  filter(CandidateName %in% c("Donald Trump", 
+                              "Kamala Harris"))
+
+all_candidates <- cleaned_data %>%
+  mutate(CandidateName = "All Candidates")
+
+combined_data <- bind_rows(
+  harris_trump_selection,
+  all_candidates
+)
+
+ggplot(combined_data, aes(x = CandidateName, 
+                          y = Percentage, 
+                          fill = CandidateName)) +
+  geom_violin(alpha = 0.5) +
+  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) +
+  labs(title = "Percentage by Candidate", 
+       x = "Candidate", 
+       y = "Percentage") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_fill_manual(values = c("grey", "salmon", "blue")) +
+  facet_wrap(~CandidateName, scales = "free_x") +
+  theme_minimal()
 ```
@@
 will update these box plots to different visuals due to box plots not showing the full dataset!
+(#Note to team: updated to violin plots to show distribution and selected only trump and harris and another one for all the candidates to show distribution of dataset)
 
@@
 # References
+
+# References

### 2024-11-02 paper/paper.qmd
Updated abstract, Introduction and reference

@@
 date-format: long
-abstract: "This study models the 2024 U.S.
-Presidential election outcomes using advanced statistical techniques and predictive modeling.
-The analysis incorporates data from various polling sources and considers key demographic and political factors to forecast both the popular vote and Electoral College results.
-The findings suggest a close race between the two main candidates, Donald Trump and Kamala Harris, with a predicted victory for Trump in the Electoral College.
-This analysis contributes to understanding how public opinion data and electoral systems interact, providing insights into electoral dynamics in contemporary U.S.
+abstract: "This paper predicts the 2024 U.S.
+Presidential election, utilizing polling data and demographic factors to predict both popular vote and Electoral College outcomes.
+The findings indicate a close race between Donald Trump and Kamala Harris, and we predict Trump secures the Electoral College due to projected victories in key swing states while Harris maintains a popular vote lead.
+The results emphasize the role of state-level dynamics, particularly in swing states, in determining election outcomes.
+The analysis contributes how public opinion data and electoral systems interact, providing insights into electoral dynamics in contemporary U.S.
 politics."
 format: pdf
+toc: true
+toc-depth: 3
 number-sections: true
@@
 
-Elections in the United States, particularly presidential elections, are pivotal moments that shape not only the nation’s political landscape but also its international relations.
-The 2024 U.S.
-Presidential election, like its predecessors, promises to be contentious and closely watched.
-As electoral outcomes have become increasingly difficult to predict due to polarized political climates and fluctuating public opinion, sophisticated statistical models are essential tools for forecasting results and understanding the dynamics at play.
+Presidential elections in the United States are pivotal moments that shape not only the nation’s political landscape but also its far-reaching consequences on a global scale.
+Republicans and Democrats held primary elections across the country earlier in 2024 that eventually filtered down to the main candidates: Donald Trump and Kamala Harris.
+As electoral outcomes have become increasingly difficult to predict due to polarized political climates and fluctuating public opinion, sophisticated statistical models are essential for forecasting results and understanding the dynamics at play.
 
-The core estimand of this study is the projected electoral outcome in terms of state-level victories and Electoral College votes for each candidate.
-The model leverages polling averages, historical voting patterns, and demographic predictors to estimate the probability of each candidate winning in individual states.
+The polarized political climate and rapidly shifting public opinion have made predicting outcomes increasingly challenging, where winning key swing states is crucial to securing victory.
+The core estimand of this study is the projected electoral outcome in terms of state-level victories and Electoral College votes for the candidate.
+The model incorporates recent polling data, historical voting patterns, and demographic predictors to estimate the probability of each candidate winning in individual states.
 The objective is to translate these probabilities into a forecast of Electoral College votes, thus determining the likely overall winner of the election.
-Given the significant influence of swing states, the model focuses on capturing their dynamics more accurately than national polling averages typically allow.
 
-The results suggest a close race, with Trump holding a slight edge in the Electoral College due to projected victories in key swing states, despite Harris maintaining a lead in national popular vote estimates.
+Our findings reveal a close race, with Trump holding a slight edge in the Electoral College due to projected victories in key swing states, despite Harris maintaining a lead in national popular vote estimates.
 Trump's path to victory hinges on critical wins in states such as Pennsylvania, Georgia, and Wisconsin, where his projected margins are narrow but statistically significant.
 Meanwhile, Harris secures strong performances in populous Democratic strongholds like California and New York, but these do not offset Trump’s advantage in less populous yet crucial battleground states.
-
-Understanding these projected outcomes is essential for several reasons.
-First, the model highlights the limitations of relying solely on national polling to predict election results, emphasizing the continued relevance of the Electoral College system in determining the presidency.
-Additionally, it sheds light on the evolving political landscape in swing states, where demographic changes and shifting voter allegiances can have outsized impacts.
+This result underscores the significant influence of swing states and the Electoral College system in determining the outcome, particularly in closely contested elections.
 These insights contribute to a broader understanding of how statistical modeling can inform political strategy and election analysis.
 
-The structure of the paper is as follows: In @sec-data, we provide outlines the data sources and variables considered, followed by the model setup in [@sec-mod] and justification.
-The results [@sec-result] presents the key findings of the analysis, with a discussion on the implications of these results for political forecasting and public opinion analysis.
-Finally, the paper concludes with an overview of potential limitations and suggestions for future research in the discussion [@sec-discussion].
+The structure of the paper is as follows: [@sec-data] outlines the data sources and variables considered, followed by the model setup in [@sec-modset] and justification in [@sec-modjust].
+The results in [@sec-result] presents the key findings of the analysis, with a discussion on the implications.
+[@sec-discussion] then discusses potential limitations and suggestions for future research.
+[@sec-appx] provides additional detailed information about the data, model and methodology.
 
@@
 The data used in this analysis comes from a combination of publicly available polling data for the 2024 U.S.
-Presidential election[@FiveThirtyEightPolls].
-The analysis leverages the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `janitor` [@janitor], `knitr` [@knitr], `dplyr` [@dplyr], `purrr` [@purrr], and `here` [@here], for data manipulation, and `ggplot2` [@ggplot] for visualization.
+Presidential election [@FiveThirtyEightPolls].
+The analysis leverages the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `janitor` [@janitor], `knitr` [@knitr], `dplyr` [@dplyr], `arrow` [@arrow], `purrr` [@purrr], and `here` [@here], for data manipulation.
+`ggplot2` [@ggplot] and `kableExtra` [@kableExtra] for visualization.
 The dataset covers various polls conducted across multiple states, capturing the support for each major candidate—Donald Trump and Kamala Harris—along with detailed attributes of the polls.
@@
 The primary measurements in the dataset include polling scores, sample sizes, and predicted percentages for each candidate.
-The data is filtered to include only polls relevant to the two primary candidates, and features like PollRecency were constructed to weigh recent polls higher.
+The data is filtered to include only polls relevant to the two primary candidates，Trump and Harris, and features like PollRecency were constructed to weigh recent polls higher.
 The data was cleaned by removing irrelevant or incomplete observations, and new variables were constructed where necessary.
@@
 
-# Model {#sec-mod}
+# Model
 
@@
 
-## Model Set-Up
+## Model Set-Up {#sec-modset}
 
@@
 
-## Model Justification
+## Model Justification {#sec-modjust}
 
@@
 
-# Appendix
+# Appendix {#sec-appx}
 
@@
 # References
-
-# References

### 2024-11-03 scripts/00-simulate_data.R
updated to use testthat and add variable interaction

@@ -4,11 +4,11 @@
 # Date: October 22nd, 2024
 # Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` and 'arrow' packages must be installed
+# Pre-requisites: The `tidyverse` package must be installed
 # Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
 # load libraries
-library(dplyr)
+library(tidyverse)
 library(arrow)
 
 # set seed for reproducability
@@ -55,6 +55,19 @@ poll_data <- data.frame(
   pct = round(runif(n, 30, 60), 2)
 )
 
+# add interactions between variables
+poll_data <- poll_data %>%
+  mutate(
+    pollscore_pct_interaction = pollscore * pct,
+    sample_size_grade_interaction = sample_size * numeric_grade,
+    sample_size_methodology_interaction = case_when(
+      methodology == "Phone" ~ sample_size * 1.1,
+      methodology == "Online" ~ sample_size * 0.9,
+      methodology == "Mixed" ~ sample_size * 1.0,
+      TRUE ~ sample_size
+    )
+  )
+
 # write the cleaned and simulated polling data to a CSV file
-write_csv(cleaned_data, here::here("data/simulated_data/simulated_data.csv"))
-write_parquet(cleaned_data, here::here("data/simulated_data/simulated_data.parquet"))
+write_csv(poll_data, here::here("data/simulated_data/simulated_data.csv"))
+write_parquet(poll_data, here::here("data/simulated_data/simulated_data.parquet"))

### 2024-11-03 scripts/01-test_simulated_data.R

@@ -5,36 +5,29 @@
 # Date: October 22nd, 2024
 # Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` and 'arrow' packages must be installed
+# Pre-requisites: The `tidyverse`, 'arrow', and `testthat` packages must be installed
 # Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
 # load libraries
 library(dplyr)
-library(purrr)
 library(arrow)
+library(testthat)
 
 # read the simulated data
 data <- read_parquet(here::here("data/simulated_data/simulated_data.parquet"))
 
-# initialize an empty list to store the results of each test (pass/fail)
+# create list to store test results
 test_results <- list()
 
 # test if the data structure is correct
-test_structure <- function(data) {
-  if (!is.data.frame(data)) {
-    return("poll_data is not a data frame.")
-  }
-  if (ncol(data) != 22) {
-    return("poll_data does not have the correct number of columns.")
-  }
-  if (nrow(data) != 100) {
-    return("poll_data does not have the correct number of rows.")
-  }
-  return("Structure test passed.")
-}
+test_results$structure <- test_that("Data structure", {
+  expect_true(is.data.frame(data), info = "poll_data is not a data frame.")
+  expect_equal(ncol(data), 25, info = "poll_data does not have the correct number of columns.")
+  expect_equal(nrow(data), 100, info = "poll_data does not have the correct number of rows.")
+})
 
 # test if the column names match the expected column names
-test_column_names <- function(data) {
+test_results$column_names <- test_that("Column names", {
   expected_colnames <- c(
     "poll_id", "pollster_id", "pollster", "sponsor_ids",
     "sponsors", "pollster_rating_id", "numeric_grade",
@@ -43,73 +36,45 @@ test_column_names <- function(data) {
     "sponsor_candidate_party", "endorsed_candidate_name",
     "endorsed_candidate_party", "sample_size",
     "population", "election_date", "stage",
-    "candidate_name", "pct"
+    "candidate_name", "pct", "pollscore_pct_interaction",
+    "sample_size_grade_interaction",
+    "sample_size_methodology_interaction"
   )
-  if (!all(colnames(data) == expected_colnames)) {
-    return("Column names do not match expected names.")
-  }
-  return("Column names test passed.")
-}
+  expect_equal(colnames(data), expected_colnames, info = "Column names do not match expected names.")
+})
 
[…]

### 2024-11-03 scripts/04-test_analysis_data.R

@@ -4,93 +4,72 @@
 # Date: October 22nd, 2024
 # Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse`, 'here', and 'arrow' packages must be installed
+# Pre-requisites: The `tidyverse`, `here`, `arrow`, and `testthat` packages must be installed.
 # Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
 # load libraries
 library(tidyverse)
 library(here)
 library(arrow)
+library(testthat)
 
 # load data
 cleaned_data <- read_parquet(here::here("data/analysis_data/president_polls_cleaned.parquet"))
 
-# define list of tests
-tests <- list(
-  # test structure of data
-  structure_test = function(data) {
-    if (!inherits(data, "tbl_df")) {
-      stop("Structure test failed: Data is not a tibble.")
-    }
-    if (ncol(data) == 0) {
-      stop("Structure test failed: Data has no columns.")
-    }
-    if (nrow(data) == 0) {
-      stop("Structure test failed: Data has no rows.")
-    }
-    cat("Structure test passed.\n")
-  },
-  # test if column names match expected names
-  column_names_test = function(data) {
-    expected_colnames <- c(
-      "poll_id", "pollster_id", "Pollster", "Party", "sponsors",
-      "NumericGrade", "PollScore", "Methodology",
-      "State", "StartDate", "EndDate",
-      "SampleSize", "Population", "ElectionDate",
-      "Stage", "CandidateName", "Percentage"
-    )
+# create list for results
+test_results <- list()
 
-    if (!all(names(data) == expected_colnames)) {
-      stop("Column names test failed: Column names do not match expected names.")
-    }
-    cat("Column names test passed.\n")
-  },
-  # test if certain columns contain NA values
-  na_values_test = function(data) {
-    if (any(is.na(data$Percentage))) {
-      stop("NA values test failed: Percentage contains NA values.")
-    }
-    if (any(is.na(data$CandidateName))) {
-      stop("NA values test failed: CandidateName contains NA values.")
-    }
-    cat("NA values test passed.\n")
-  },
-  # test if any numeric grades or pollscores are out of expected range
-  numeric_grade_pollscore_test = function(data) {
-    if (any(data$NumericGrade < 0 | data$NumericGrade > 3, na.rm = TRUE)) {
-      stop("Numeric grade test failed: NumericGrade is out of expected range.")
-    }
[…]

### 2024-11-03 paper/paper.qmd
Updated measurement and syllabus

@@
 abstract: "This paper predicts the 2024 U.S.
-Presidential election, utilizing polling data and demographic factors to predict both popular vote and Electoral College outcomes.
-The findings indicate a close race between Donald Trump and Kamala Harris, and we predict Trump secures the Electoral College due to projected victories in key swing states while Harris maintains a popular vote lead.
-The results emphasize the role of state-level dynamics, particularly in swing states, in determining election outcomes.
-The analysis contributes how public opinion data and electoral systems interact, providing insights into electoral dynamics in contemporary U.S.
-politics."
+Presidential election.
+We utilizes linear regression model deploying polling data and demographic factors to predict both popular vote and Electoral College outcomes.
+The findings indicate a close race between Donald Trump and Kamala Harris, and we predict Trump will secure the Electoral College due to projected victories in key swing states while Harris maintains a popular vote lead.
+The results emphasize the role of state-level dynamics, particularly in swing states, in determining election outcomes."
 format: pdf
@@
 
-Presidential elections in the United States are pivotal moments that shape not only the nation’s political landscape but also its far-reaching consequences on a global scale.
+The U.S.
+presidential election is a pivotal moment not only shapes domestic political landscape but also far-reaching its consequences on a global scale.
 Republicans and Democrats held primary elections across the country earlier in 2024 that eventually filtered down to the main candidates: Donald Trump and Kamala Harris.
-As electoral outcomes have become increasingly difficult to predict due to polarized political climates and fluctuating public opinion, sophisticated statistical models are essential for forecasting results and understanding the dynamics at play.
+The polarized political climate and rapidly shifting public opinion have made predicting outcomes increasingly challenging, where winning key swing states is crucial to securing victory.
 
-The polarized political climate and rapidly shifting public opinion have made predicting outcomes increasingly challenging, where winning key swing states is crucial to securing victory.
-The core estimand of this study is the projected electoral outcome in terms of state-level victories and Electoral College votes for the candidate.
-The model incorporates recent polling data, historical voting patterns, and demographic predictors to estimate the probability of each candidate winning in individual states.
-The objective is to translate these probabilities into a forecast of Electoral College votes, thus determining the likely overall winner of the election.
+The main estimand is the projected support percentage and each candidate's winning probability in terms of state-level victories and Electoral College votes.
+This paper utilizes linear regression model incorporates numeric value of the credibility of pollsters, number of respondents, poll's recency to estimate the probability of each candidate winning in individual states.
+The objective is to translate these probabilities into a forecast the overall winner of the election.
+Our findings reveal a close race, with Trump holding a slight edge in the Electoral College due to projected victories in key swing states, despite Harris maintaining a lead in national popular vote estimates.
 
-Our findings reveal a close race, with Trump holding a slight edge in the Electoral College due to projected victories in key swing states, despite Harris maintaining a lead in national popular vote estimates.
 Trump's path to victory hinges on critical wins in states such as Pennsylvania, Georgia, and Wisconsin, where his projected margins are narrow but statistically significant.
@@
 
-The primary measurements in the dataset include polling scores, sample sizes, and predicted percentages for each candidate.
-The data is filtered to include only polls relevant to the two primary candidates，Trump and Harris, and features like PollRecency were constructed to weigh recent polls higher.
+The primary measurement process reflects the transformation of raw polling data into structured entries to actionable insights for election forecasting.
+We first select and rename key variables from raw data to focus on relevant information, such as pollster, pollscore, state, polling dates, sample size, and candidate support percentage.
+These variables represent essential poll characteristics that capture public sentiment in response to real-world campaign dynamics and demographic differences across states.
+
+To address instances where the original data referenced Joe Biden instead of Kamala Harris, we assumed the entries for Biden will goes to Harris in consequences of Democratic Party.
+This approach is supported by "Harris’s truncated presidential campaign relies heavily upon President Biden’s policy framework and she eaddress the campaign’s weaknesses and win over critical voter demographics [@RealInstitutoElcano].
+Thus, substituting Biden’s polling data with Harris’s ensures continuity in demographic appeal without altering established polling trends.
+Then we filter to include only polls relevant to the two candidates，Trump and Harris, and features like PollRecency were constructed to weigh recent polls higher.
 The data was cleaned by removing irrelevant or incomplete observations, and new variables were constructed where necessary.
-
-## Outcome Variables
-
-The main outcome variable of interest is the Percentage of support each candidate has in a given poll, which represents the proportion of respondents who favor one candidate over the other.
-This variable is crucial for estimating both the popular vote and Electoral College outcomes.
 
@@
 
+## Outcome Variables
+
+The main outcome variable of interest is the Percentage of support each candidate has in a given poll, which represents the proportion of respondents who favor one candidate over the other.
+This variable is crucial for estimating both the popular vote and Electoral College outcomes.
+
 # Model
@@
 
-This paper utilized a linear model: $$Percentage_i = \beta_0 + \beta_1 Pollscore_i + \beta_2 SampleSize + \beta_3 Numeric Grade_i + \beta_4 State_i + \beta_5PollRecency + \epsilon_i$$ where, Percentage$_i$ represents the predicted support percentage for a candidate in a given poll.
+This paper utilizes linear regression model: $$Percentage_i = \beta_0 + \beta_1 Pollscore_i + \beta_2 SampleSize + \beta_3 Numeric Grade_i + \beta_4 State_i + \beta_5PollRecency + \epsilon_i$$ where, Percentage$_i$ represents the predicted support percentage for a candidate in a given poll.
 

### 2024-11-03 paper/paper.qmd
updated table names and captions

--- 2024-11-02 paper/paper.qmd
@@
 ```{r}
+#| label: tbl-popularvote
+#| tbl-cap: The percentage of support Harris and Trump receive overall
 #| echo: false
@@
 ```{r}
+#| label: tbl-electoral
+#| tbl-cap: The number of electoral college votes Harris vs Trump recieves
 #| echo: false
@@
 ```{r}
+#| label: tbl-statewinners
+#| tbl-cap: The winners of each state, which determines who gets the electoral college votes for that state.
 #| echo: false
@@
 #| label: tbl-sum
-#| fig-cap: [Need Edit]
+#| fig-cap: summary statistics for the cleaned data
 #| echo: false
@@
 #| label: fig-predictedscoremap
-#| fig-cap: "Mapped Model-predicted outsom for 2024 elections by State"
+#| fig-cap: "Mapped Model-predicted outcomes for 2024 elections by State"
 #| echo: false
@@
 
+## Pollster Methodology
+
 ## Idealized Methodology

### 2024-11-03 paper/paper.qmd
Edited model descriptions

@@
 To predict the outcome of the 2024 U.S.
-Presidential election, we developed two linear regression models: one for Donald Trump and one for Kamala Harris.
-These models aim to estimate the percentage of support for each candidate based on the aforementioned predictor variables.
+Presidential election, we developed several linear regression models: one for the popular vote and one for the electoral college.
+These models aim to estimate the percentage of support for each candidate, filtered to Harris and Trump, based on the aforementioned predictor variables.
+This is done both nationally at a popular vote level, but also by state to determine the distribution of electoral votes.
 

### 2024-11-03 paper/paper.qmd
Add files via upload

--- 2024-11-03 paper/paper.qmd
@@
 
+## Outcome Variables
+
+The main outcome variable of interest is the Percentage of support each candidate has in a given poll, which represents the proportion of respondents who favor one candidate over the other.
+This variable is crucial for estimating both the popular vote and Electoral College outcomes.
+
 ## Predictor Variables
@@
 CandidateName: The name of the candidate being polled (either Donald Trump or Kamala Harris).
-
-## Outcome Variables
-
-The main outcome variable of interest is the Percentage of support each candidate has in a given poll, which represents the proportion of respondents who favor one candidate over the other.
-This variable is crucial for estimating both the popular vote and Electoral College outcomes.
 

### 2024-11-03 paper/paper.qmd
update pollscore variable definition
Completed idealized poll + pollster deep dive

--- 2024-11-03 paper/paper.qmd
@@
 PollScore: A numeric value that reflects the credibility of the pollster.
+Stands for "Predictive Optimization of Latent skill Level in Surveys, Considering Overall Record, Empirically." The error and bias we can attribute to a pollster.
+Negative numbers are better.
 
@@
 
-## Additional Data Details
+## Pollster Methodology
 
-## Model Details {#sec-model-details}
+The New York Times/Siena College Poll was ranked the most accurate political pollster in the US following the 2022 midterm elections.
+They received the max 538 rating of 3 stars, one of only four pollsters to do so.
+They also had a pollscore of -1.5, meaning they had the least amount of error and bias attributed to them of the 282 pollsters ranked by 538.
 
-## Pollster Methodology
+The Times/Siena Poll targets registered voters in the United States, emphasizing those in battleground states during presidential elections.
+Since the US decides it's President based on the Electoral College rather than the national popular vote, this is an important focus in order to correctly determine election outcomes.
+The sampling frame is derived from a voter file which containing lists of 200 million registered voters as well as their demographic information and prior voting behavior data.
+
+Each poll typically surveys around 1,000 respondents, which is deemed sufficient for achieving a margin of sampling error of approximately 3-4 percentage points.
+The Times found that tripling the sample size would only reduce the margin of error by a percent or two.
+This sample size allows for a balance between representativeness and efficiency.
+
+Respondents are recruited through live phone interviews conducted by trained interviewers from call centers in various states.
+The polling effort includes both landline and cellphone calls, with over 90% of respondents being reached via cellphones.
+
+Times/Siena uses a stratified random sampling technique, where the sample is divided into strata based on demographics such as age, race, gender, and party affiliation.
+This helps to ensure proportional representation from each group.
+Stratification helps in achieving a sample that mirrors the voting population, minimizing biases related to demographic variables (i.e.
+uneducated voters are less likely to respond to polls).
+However, it faces limitations as stratification increases the complexity of the sampling process, and if strata are not adequately represented in the sample, the results may still skew.
+Times/Siena combats this by applying weighting adjustments post-survey to correct for any demographic discrepancies.
+
+Given the declining response rates typical of phone surveys, the Times/Siena Poll faces challenges in reaching respondents.
+Approximately 2% of those contacted ultimately participate.
+This pollster uses follow-up calls to make multiple attempts to reach individuals who initially do not respond.
+In 2022, they conducted an experiment to find out if responses were skewed by non-response bias.
+This was done via a mail survey that offered a reward of up to \$25 for a response.
+This saw a 30% response rate, which is significantly larger than the 2% response rate Times/Siena gets over phone.
+Although there were differences between the types of respondents you get between the two methodologies, there was not a meaningful difference in who these respondants claim they would vote for.
+
+The questionnaires used by Times/Siena are designed to capture a broad range of opinions while remaining concise.
+Interviews are kept under 15 minutes to reduce survey fatigue and maximize response rates.
+This pollster prioritizes making sure every viewpoint is accurately represented in order to make the questions fair and ensure that respondents are not being swayed in a particular direction.
+This means the questions must remain objective and clear.
+However, a survey with limited response options and questions may mean a respondent's view is not captured in its full breadth.
+It also means that more nuanced opinions may be overlooked in favor of broader trends.
+
+The New York Times/Siena College Poll is rated the most accurate pollster in the US for a reason - it has a well-defined methodology with a fleshed out population frame, sampling approach, and contingencies for non-response handling.
 
@@
 
-This is an outline. I will turn this into paragraph style this week.
+The idealized methodology for forecasting the US presidential election targets eligible registered voters from all 50 states and Washington D.C., ensuring representation from each state in order to create an Electoral College forecast, with particular emphasis on swing (battleground) states.
 
-**Target Population:** The target population is eligible US voters from all 50 states and Washington D.C. –\> representation from every state for an Electoral College forecast.
+The sampling method will utilize stratified random sampling, which stratifies the sample by state.
+Voters will be further stratified based on factors such as age, gender, race/ethnicity, income, and political affiliation.
+In swing states, we will ensure oversampling to capture more granular data and reduce our margin of error.
 
-**Sampling Method**
+To maintain a national margin of error of ±4%, the total sample size will be 1,000 respondents per state.
+Swing states will be oversampled, with over 2,000 respondents to achieve a margin of error of ±2%.
+Given an estimated response rate of 2%, we aim to recruit approximately 50,000 potential respondents per state, with a target of 100,000 in battleground states.
 
-\- Stratified Random Sampling: The sample will be stratified by state, ensuring each state is proportionally represented in terms of population size and key voter demographics (age, gender, race/ethnicity, income, political affiliation).
+To recruit respondents, we will utilize online panel providers like YouGov, Ipsos, and Dynata to obtain geographically and demographically targeted participants.
+Additionally, we will conduct state-specific phone surveys, employing interactive voice response (IVR) systems and live calls to reach older and less tech-savvy voters, particularly in states with a higher proportion of rural populations.
+To incentivize participation, all respondents will be entered into state-level sweepstakes, with one winner selected per state.
 
-\- Key Stratification Variables: Within each state, voters will be further stratified based on party affiliation, urban/rural distribution, and voting history.
+Data collection will be conducted on platforms such as Google Forms.
+Telephone surveys will follow the layout and workflow of the Google Form survey.
+The survey is designed to take approximately five to ten minutes to complete in order to minimize drop-off rates.
+It will include national-level questions, covering voting intent, candidate preference, key issues, and voter perceptions.
+After collection, the data will be weighted by state and demographic factors to ensure accuracy, utilizing the latest Census data and state voting records.
 
-\- For swing states, we will ensure oversampling to capture more granular data.
+For data validation, we will implement several measures, including completion time checks to identify speeding responses, IP tracking to prevent duplicate entries, and reCAPTCHA to guard against bots.
+Respondents will provide their state of residence and ZIP code for additional verification.
+Furthermore, the data will be weighted according to state demographics and previous election turnout data, utilizing information from Census and state voting records.
+Special attention will be paid to swing states to ensure no key voter segments are underrepresented.
 
-**Sample Size**
+In terms of budget allocation, we will designate funds as follows:
 
-\- The total sample size will be 15,000 respondents –\> ensures state-level representation while maintaining a national margin of error of ±1%.
+1.  Survey Development (\$7,000):
+    -   The \$7,000 covers the costs associated with designing the survey, including the creation of questions, structuring the survey for clarity and flow, and ensuring that it meets methodological standards.
+It may also include hiring experts in survey design or consulting with statisticians to validate the survey framework and ensure questions are unbiased.
+Additionally, while Google Forms is free, this budget may account for platform costs if needed.
+2.  Accurate Sampling (\$10,000):
+    -   The sampling budget is allocated for the costs associated with selecting a representative sample of voters.
+This includes expenses related to stratified random sampling, where the sample is segmented by various demographics and state representation.
+In order to ensure each demographic share is accurately represented, we may need to purchase access to demographic data or employ statisticians to ensure that the sampling aligns with the population of each state.
+3.  Recruiting Respondents (\$45,000):
+    -   The majority of the budget is dedicated to recruiting respondents, which is necessary for achieving a large and demographically representative sample.
+This cost will cover payments to online panel providers which charge fees for access to their respondent pools.
+Additionally, this budget will cover costs related to conducting state-specific phone surveys and IVR calls, such as hiring call center staff or IVR technology.
+4.  Respondent Incentives (\$28,000):
+    -   To encourage participation, we plan to run state-level sweepstakes, where respondents have a chance to win prizes for completing the survey.
+A \$500 prize will be awarded to one winner in each state (+ Washington DC).
+The remaining budget will cover administrative expenses associated with running the sweepstakes.
+5.  Data Analysis and Reporting (\$10,000):
+    -   This portion of the budget is reserved for the costs related to analyzing the collected data and reporting the findings.
+This may include hiring data analysts or consultants who specialize in statistical analysis and survey data interpretation.
+The funds will also cover software tools necessary for data analysis, as well as the creation of visualizations and reports to communicate the results effectively.
 
-\- As mentioned above, swing states will have an additional focus, with 1,000+ respondents per state (depending on population size).
-
-\- Electoral vote-rich states (e.g., California, Texas, Florida) will be proportionally represented to reflect their impact on the election outcome.
-
-\- With an estimated response rate of 10%, we aim to recruit 150,000 potential respondents across states.
-
-**Recruiting Respondents**
-
-\- Panel Providers: Utilize online panel providers (such as YouGov, Ipsos, or Dynata) for geographically and demographically targeted respondents by state.
-
-\- Social Media Ads: Conduct state-targeted ads on platforms like Facebook and Instagram\*for broad geographic reach
-
-\- State-Specific Phone Surveys: To reach older and less tech-savvy voters, particularly in states with higher proportions of rural populations, we will use IVR phone surveys and live calls.
-
-\- Respondents will be entered into state-level sweepstakes, with one winner per state
-
-**Data Collection**
-
-\- Surveys will be conducted on platforms like Qualtrics or SurveyMonkey, with branching logic for state-specific questions.
-Telephone surveys for older demographics will be integrated.
-The mock survey will be conducted using Google Forms for the purposes of accessibility.
-
-\- The survey will take approximately 5 minutes in order to minimize drop-off rates
-
-\- The survey will include state-specific questions and national-level questions, such as voting intent, candidate preference, key issues, and voter perception.
-
-\- Weighting by State: Data will be weighted post-collection to ensure state-level accuracy, considering the latest Census data and state voting records.
-
-**Data Validation**
-
-\- Validate completion time (speeding checks), IP tracking (to prevent duplicate responses), and reCAPTCHA to prevent bots.
-
-\- Respondents will be asked their state of residence, and ZIP code for additional verification.
-
-\- Data will be weighted by state demographics and previous election turnout using Census and state voting data.
-Swing states will receive additional attention to ensure no underrepresentation of key voter segments.
-
-include budget allocation?
-
-**Survey Structure (will be migrated to google form)**
-
-Introduction
-
-\- Brief description of the survey’s purpose (US Presidential Election Forecast with a focus on state-level dynamics).
-
-\- Assurance of confidentiality and anonymity, with a clear outline of data use.
-
-\- Contact details for survey-related inquiries.
-
-Questions
-
-1\. Demographics:
-
-\- State of Residence (Dropdown list of all 50 states and D.C.)
-
-\- Age, gender, race/ethnicity, income level, education, party affiliation (Republican, Democrat, Independent, Other).
-
-2\. Voter Registration:
-
-\- Are you registered to vote in the upcoming election? (Yes/No)
-
-\- How likely are you to vote in the upcoming election? (Likert scale)
-
-3\. Voting Intent:
-
-\- If the election were held today, who would you vote for? (List of candidates specific to the respondent’s state, based on ballot access).
-
-\- How certain are you of your choice? (Likert scale)
-
-4\. Key Issues:
-
-\- What are the most important issues influencing your vote? (Rank the top 3 issues: Economy, Healthcare, Immigration, Climate Change, National Security, Social Issues, etc.)
-
-5\. State-Specific Issues:
-
-\- What state-level issues are important to you? (Open-ended question, with suggestions like state economy, healthcare access, local infrastructure, etc.)\
-
-6\. Perceptions of Candidates:
-
-\- Rate the following candidates on trustworthiness, competence, leadership, etc. (Likert scale, customized for state-level candidates where applicable).
-
-7\. Previous Voting Behavior:
-
-\- Did you vote in the last presidential election? (Yes/No)
-
-\- If yes, who did you vote for? (Open response, with drop-down of major candidates for validation)
-
-Closing
-
-\- Thank You: Thank the respondents for their time and participation, reminding them of the sweepstakes.
+An example of the survey can be found at this link: <https://forms.gle/igzK683kHLYytGvu8>
 

### 2024-11-04 paper/paper.qmd
Added citations to appendix

@@
 
-\[WILL CHANGE THE FORMAT OF THESE GRAPHS AND CAPTION/DESCRIPTION WILL BE ADDED\]
-
 ```{r}
@@
 #| label: fig-samplesize
-#| fig-cap: [Need Edit]
+#| fig-cap: The distribution of sample sizes used in each poll
 #| echo: false
@@
 #| label: fig-percentage
-#| fig-cap: [Need Edit]
+#| fig-cap: The distribution of what percentage of support a candidate recieves in a poll
 #| echo: false
@@
 #| label: fig-candidatepercentage
-#| fig-cap: [Need Edit]
+#| fig-cap: violin plot displaying the percentage of support Harris and Trump receive compared to all candidates
 #| echo: false
@@
 
-will update these box plots to different visuals due to box plots not showing the full dataset!
-(#Note to team: updated to violin plots to show distribution and selected only trump and harris and another one for all the candidates to show distribution of dataset)
-
 ```{r}
 #| label: fig-statepercentage
-#| fig-cap: [Need Edit]
+#| fig-cap: The percentage of support seen by state
 #| echo: false
@@
   labs(title = "Percentage by State", x = "State", y = "Percentage") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1))
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 6)) 
 ```
@@
 #| label: fig-corrmatrix
-#| fig-cap: [Need Edit]
+#| fig-cap: correlation between PollScore, the sample size of the poll, and the percentage of support for a candidate
 #| echo: false
@@
 
-The New York Times/Siena College Poll was ranked the most accurate political pollster in the US following the 2022 midterm elections.
+The New York Times/Siena College Poll was ranked the most accurate political pollster in the US following the 2022 midterm elections [@PollsterRating].
 They received the max 538 rating of 3 stars, one of only four pollsters to do so.
-They also had a pollscore of -1.5, meaning they had the least amount of error and bias attributed to them of the 282 pollsters ranked by 538.
+They also had a pollscore of -1.5, meaning they had the least amount of error and bias attributed to them of the 282 pollsters ranked by 538 [@PollsterRating].
 
@@
 Since the US decides it's President based on the Electoral College rather than the national popular vote, this is an important focus in order to correctly determine election outcomes.
-The sampling frame is derived from a voter file which containing lists of 200 million registered voters as well as their demographic information and prior voting behavior data.
+The sampling frame is derived from a voter file which containing lists of 200 million registered voters as well as their demographic information and prior voting behavior data [@TimesSienaPoll].
 
@@
 The Times found that tripling the sample size would only reduce the margin of error by a percent or two.
-This sample size allows for a balance between representativeness and efficiency.
+This sample size allows for a balance between representation and efficiency [@TimesSienaPoll].
 
 Respondents are recruited through live phone interviews conducted by trained interviewers from call centers in various states.
-The polling effort includes both landline and cellphone calls, with over 90% of respondents being reached via cellphones.
+The polling effort includes both landline and cellphone calls, with over 90% of respondents being reached via cellphones [@TimesSienaPoll].
 
@@
 However, it faces limitations as stratification increases the complexity of the sampling process, and if strata are not adequately represented in the sample, the results may still skew.
-Times/Siena combats this by applying weighting adjustments post-survey to correct for any demographic discrepancies.
+Times/Siena combats this by applying weighting adjustments post-survey to correct for any demographic discrepancies [@TimesSienaPoll].
 
@@
 Approximately 2% of those contacted ultimately participate.
-This pollster uses follow-up calls to make multiple attempts to reach individuals who initially do not respond.
-In 2022, they conducted an experiment to find out if responses were skewed by non-response bias.
+This pollster uses follow-up calls to make multiple attempts to reach individuals who initially do not respond [@TimesSienaPoll].
+In 2022, they conducted an experiment to find out if responses were skewed by non-response bias [@NonResponseBias].
 This was done via a mail survey that offered a reward of up to \$25 for a response.
 This saw a 30% response rate, which is significantly larger than the 2% response rate Times/Siena gets over phone.
-Although there were differences between the types of respondents you get between the two methodologies, there was not a meaningful difference in who these respondants claim they would vote for.
+Although there were differences between the types of respondents you get between the two methodologies, there was not a meaningful difference in who these respondants claim they would vote for [@NonResponseBias].
 
 The questionnaires used by Times/Siena are designed to capture a broad range of opinions while remaining concise.
-Interviews are kept under 15 minutes to reduce survey fatigue and maximize response rates.
+Interviews are kept under 15 minutes to reduce survey fatigue and maximize response rates [@TimesSienaPoll].
 This pollster prioritizes making sure every viewpoint is accurately represented in order to make the questions fair and ensure that respondents are not being swayed in a particular direction.
@@
 
-To maintain a national margin of error of ±4%, the total sample size will be 1,000 respondents per state.
+To maintain a national margin of error of ±4%, based on the experiment conducted by Times/Siena, the total sample size will be 1,000 respondents per state [@TimesSienaPoll].
 Swing states will be oversampled, with over 2,000 respondents to achieve a margin of error of ±2%.
@@
 
-1.  Survey Development (\$7,000):
+1.  Survey Development: \$7,000
     -   The \$7,000 covers the costs associated with designing the survey, including the creation of questions, structuring the survey for clarity and flow, and ensuring that it meets methodological standards.
@@
 Additionally, while Google Forms is free, this budget may account for platform costs if needed.
-2.  Accurate Sampling (\$10,000):
+2.  Accurate Sampling: \$10,000
     -   The sampling budget is allocated for the costs associated with selecting a representative sample of voters.
@@
 In order to ensure each demographic share is accurately represented, we may need to purchase access to demographic data or employ statisticians to ensure that the sampling aligns with the population of each state.
-3.  Recruiting Respondents (\$45,000):
+3.  Recruiting Respondents: \$45,000
     -   The majority of the budget is dedicated to recruiting respondents, which is necessary for achieving a large and demographically representative sample.
@@
 Additionally, this budget will cover costs related to conducting state-specific phone surveys and IVR calls, such as hiring call center staff or IVR technology.
-4.  Respondent Incentives (\$28,000):
+4.  Respondent Incentives: \$28,000
     -   To encourage participation, we plan to run state-level sweepstakes, where respondents have a chance to win prizes for completing the survey.
@@
 The remaining budget will cover administrative expenses associated with running the sweepstakes.
-5.  Data Analysis and Reporting (\$10,000):
+5.  Data Analysis and Reporting: \$10,000
     -   This portion of the budget is reserved for the costs related to analyzing the collected data and reporting the findings.

### 2024-11-04 paper/paper.qmd
Updated model section

@@
 To predict the outcome of the 2024 U.S.
-Presidential election, we developed two linear regression models: one for Donald Trump and one for Kamala Harris.
+Presidential election, we developed two linear regression models: one for Kamala Harris and one for Donald Trump.
 These models aim to estimate the percentage of support for each candidate based on the aforementioned predictor variables.
@@
 
-This paper utilizes linear regression model: $$Percentage_i = \beta_0 + \beta_1 Pollscore_i + \beta_2 SampleSize + \beta_3 Numeric Grade_i + \beta_4 State_i + \beta_5PollRecency + \epsilon_i$$ where, Percentage$_i$ represents the predicted support percentage for a candidate in a given poll.
+This paper utilizes linear regression models:
 
-$\beta_i$ are the coefficients of the model that quantify the impact of each predictor.
+$$Percentage_H = \beta_0 + \beta_1 Pollscore_i + \beta_2 SampleSize + \beta_3 Numeric Grade_i + \beta_4 State_i + \beta_5PollRecency + \epsilon_i$$
 
-$\epsilon_i$ is the error term.
+$$Percentage_T = \beta_0 + \beta_1 Pollscore_i + \beta_2 SampleSize + \beta_3 Numeric Grade_i + \beta_4 State_i + \beta_5PollRecency + \epsilon_i$$In this model, $Percentage$ represents the predicted support percentage for a candidate in a given poll.
+$Percentage_H$ and $Percentage_T$ represent the predicted support for Harris and Trump respectively.
+Each predictor variable is carefully selected to capture critical factors influencing voter support, with coefficients $\beta_i$ representing the strength of each factor's impact.
+The error term $\epsilon_i$ accounts for random variability not explained by the model.
 
-The results from these models are aggregated to compute the predicted percentage of support for each candidate in each state.
-Based on these percentages, we forecast the likely winner of the state’s electoral votes.
+For a more granular analysis, separate linear regression models were constructed for each of the two primary candidates, Harris and Trump, to predict their state-by-state poll percentages.
+These models allow for a detailed view of each candidate’s performance across states.
+Based on these percentages, we forecast the likely winner of the popular vote as well as the state’s electoral votes.
+These models filter for $CandidateName$ since the model is candidate-specific.
+
+As defined in [@sec-data], each predictor variable has been chosen to reflect characteristics that influence election polling.
 
@@
 
-This model was chosen for its balance between simplicity and predictive power.
+While other complex models could be appropriate in this instance, these models were chosen for its balance between simplicity and predictive power.
 Linear regression allows us to capture relationships between polling quality, sample size, and recency, while still providing interpretable coefficients.
 Given the nature of polling data, this approach is both appropriate and commonly used in election forecasting.
-The model was validated using out-of-sample testing and Root Mean Square Error to ensure predictive accuracy.
+A linear model allows us to directly interpret the influence of each predictor on voter support.
+
+## Assumptions and Limitations
+
+This model operates under several assumptions, including that relationships between predictors and outcomes are linear, errors are normally distributed, and predictors are not excessively collinear.
+Potential limitations include the fact that linear regression may oversimplify relationships in polling data, especially if interactions between predictors exist.
+Furthermore, as the model relies on historical polling data, it assumes that past trends are indicative of future behavior, which may not hold true if significant political events alter voter preferences unexpectedly, as discussed in [@sec-discussion].
+
+## Model Validation
+
+The model was implemented using R, with out-of-sample testing and Root Mean Square Error (RMSE) calculations applied to assess predictive accuracy as predicted in @tbl-rmse.
+By splitting the data into training and test sets, we evaluated how well the model performs on unseen data.
+This approach helps ensure that the model generalizes effectively to new data and avoids overfitting.
+
+```{r}
+#| label: tbl-rmse
+#| tbl-cap: The RMSE for each model
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+rmse_results <- read_parquet(here::here("data/analysis_data/rmse_results.parquet"))
+
+rmse_results %>%
+  kable(caption = "RMSE Results for Candidate Models",
+        col.names = c("Model", "Root Mean Squared Error (RMSE)"),
+        digits = 2) %>%
+  kable_styling(latex_options = c("striped", "scale_down"))
+
+```
 
@@
 
-cleaned_data <- read_parquet(here::here("data/analysis_data/model_data.parquet"))
-lm_model <- readRDS(here::here("models/lm_model.rds"))
+cleaned_data <- read_parquet(here::here("data/analysis_data/electoral_data.parquet"))
+lm_model_trump <- readRDS(here::here("models/lm_model_trump.rds"))
+lm_model_harris <- readRDS(here::here("models/lm_model_harris.rds"))
 
 cleaned_data <- cleaned_data %>%
-  mutate(PredictedPercentage = predict(lm_model, newdata = cleaned_data))
+  mutate(PredictedTrumpPercentage = predict(lm_model_trump, newdata = cleaned_data),
+         PredictedHarrisPercentage = predict(lm_model_harris, newdata = cleaned_data))
 
 results <- cleaned_data %>%
-  group_by(CandidateName) %>%
-  summarise(AveragePredictedPercentage = mean(PredictedPercentage, na.rm = TRUE)) %>%
-  arrange(desc(AveragePredictedPercentage)) %>%
-  filter(AveragePredictedPercentage > 0)
+  summarise(
+    NationalTrumpAverage = mean(PredictedTrumpPercentage, na.rm = TRUE),
+    NationalHarrisAverage = mean(PredictedHarrisPercentage, na.rm = TRUE)
+  )
 
-total_percentage <- sum(results$AveragePredictedPercentage, na.rm = TRUE)
-
+total_percentage <- sum(results$NationalTrumpAverage, results$NationalHarrisAverage, na.rm = TRUE)
 results <- results %>%
-  mutate(NormalizedPercentage = (AveragePredictedPercentage / total_percentage) * 100)
+  mutate(
+    NormalizedTrumpPercentage = (NationalTrumpAverage / total_percentage) * 100,
+    NormalizedHarrisPercentage = (NationalHarrisAverage / total_percentage) * 100
+  )
 
 results %>%
+  select(NationalTrumpAverage, NormalizedTrumpPercentage, NationalHarrisAverage, NormalizedHarrisPercentage) %>%
   kable(
-    caption = "Predicted Percentage Vote for Donald Trump and Kamala Harris",
-    col.names = c("Candidate Name", "Average Predicted Percentage", "Normalized Percentage"),
+    caption = "Predicted National Popular Vote for Donald Trump and Kamala Harris",
+    col.names = c("Trump Average Predicted %", "Trump Normalized %", "Harris Average Predicted %", "Harris Normalized %"),
     digits = 2,
@@
   kable_styling(latex_options = c("striped", "scale_down"))
+
 ```
@@
     caption = "Predicted Number of Electoral Votes for Donald Trump and Kamala Harris",
-    col.names = c("Trumps Total Electoral Votes", "Harriss Total Electoral Votes"),
+    col.names = c("Trumps Total Electoral Votes", "Harris Total Electoral Votes"),
     digits = 2,
@@
 
-The model results (see Appendix for details) show a clear distribution of support for both candidates across various states.
+The model results show a clear distribution of support for both candidates across various states.
 The predicted percentages of support were normalized to ensure they summed up to 100%.

### 2024-11-04 scripts/06-model_data.R
Updated models and added validation

@@ -1,10 +1,10 @@
 #### Preamble ####
-# Purpose: Run a linear regression model on president polls data
+# Purpose: Run a linear regression model to predict the electoral college
 # Author: Sophia Brothers
 # Date: October 22nd, 2024
 # Contact: sophia.brothers@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse`, 'here', and 'arrow' packages must be installed
+# Pre-requisites: The `tidyverse` and 'here' packages must be installed
 # Any other information needed? Make sure you are in the `US_Election_2024` rproj
 
 # load libraries
@@ -12,10 +12,8 @@ library(tidyverse)
 library(here)
 library(arrow)
 
-# load data
-cleaned_data <- read_parquet(here::here("data/analysis_data/president_polls_cleaned.parquet"))
-
 # convert some columns to the appropriate data types
+cleaned_data <- read_parquet(here::here("data/analysis_data/president_polls_cleaned.parquet"))
 cleaned_data <- cleaned_data %>%
   mutate(
     CandidateName = as.factor(CandidateName),
@@ -36,19 +34,21 @@ cleaned_data <- cleaned_data %>%
 
 # only consider polls for Trump or Harris
 cleaned_data <- cleaned_data %>%
-  filter(CandidateName %in% c("Donald Trump", "Kamala Harris"))
+  filter(CandidateName %in% c("Donald Trump", "Kamala Harris")) %>%
+  select(State, PollScore, SampleSize, NumericGrade, PollRecency, Percentage, CandidateName)
 
-# linear regression model that predicts support percentages
-lm_model <- lm(
-  Percentage ~ PollScore + SampleSize + NumericGrade + State +
-    PollRecency + CandidateName,
-  data = cleaned_data
+# linear model for Donald Trump's poll percentage based on various predictors
+lm_model_trump <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State,
+                     data = cleaned_data %>% filter(CandidateName == "Donald Trump")
 )
 
-# summary of model
-summary(lm_model)
+# linear model for Kamala Harris' poll percentage based on various predictors
+lm_model_harris <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State,
+                      data = cleaned_data %>% filter(CandidateName == "Kamala Harris")
+)
 
[…]

### 2024-11-04 scripts/07-validation.R

@@ -1,0 +1,67 @@
+#### Preamble ####
+# Purpose: Validate the regression models through out of sample testing and RMSE
+# Author: Sophia Brothers
+# Date: October 22nd, 2024
+# Contact: sophia.brothers@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse`, 'caret', and 'here' packages must be installed
+# Any other information needed? Make sure you are in the `US_Election_2024` rproj
+
+
+# load libraries
+library(tidyverse)
+library(here)
+library(caret)
+
+# load data
+cleaned_data <- read_parquet(here::here("data/analysis_data/president_polls_cleaned.parquet"))
+
+# data processing
+cleaned_data <- cleaned_data %>%
+  mutate(
+    CandidateName = as.factor(CandidateName),
+    State = as.factor(State),
+    Pollster = as.factor(Pollster),
+    StartDate = as.Date(StartDate),
+    EndDate = as.Date(EndDate),
+    Party = as.factor(Party),
+    Methodology = as.factor(Methodology),
+    NumericGrade = as.numeric(NumericGrade),
+    SampleSize = as.numeric(SampleSize)
+  ) %>%
+  mutate(PollRecency = as.numeric(Sys.Date() - StartDate)) %>%
+  filter(CandidateName %in% c("Donald Trump", "Kamala Harris"))
+
+# split the data into training (70%) and testing (30%) sets
+set.seed(123)  # Set seed for reproducibility
+train_index <- createDataPartition(cleaned_data$CandidateName, p = 0.7, list = FALSE)
+train_data <- cleaned_data[train_index, ]
+test_data <- cleaned_data[-train_index, ]
+
+# fit linear regression models on the training set
+lm_model_trump <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State,
+                     data = train_data %>% filter(CandidateName == "Donald Trump"))
+
+lm_model_harris <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State,
+                      data = train_data %>% filter(CandidateName == "Kamala Harris"))
+
+# make predictions on the testing set for each model
+predictions_trump <- predict(lm_model_trump, newdata = test_data %>% filter(CandidateName == "Donald Trump"))
[…]

### 2024-11-04 scripts/07-electoral_data.R
Delete duplicate file

@@ -1,54 +1,0 @@
-#### Preamble ####
-# Purpose: Run a linear regression model to predict the electoral college
-# Author: Sophia Brothers
-# Date: October 22nd, 2024
-# Contact: sophia.brothers@mail.utoronto.ca
-# License: MIT
-# Pre-requisites: The `tidyverse`, 'here', and 'arrow' packages must be installed
-# Any other information needed? Make sure you are in the `US_Election_2024` rproj
-
-# load libraries
-library(tidyverse)
-library(here)
-library(arrow)
-
-# convert some columns to the appropriate data types
-cleaned_data <- read_parquet(here::here("data/analysis_data/president_polls_cleaned.parquet"))
-cleaned_data <- cleaned_data %>%
-  mutate(
-    CandidateName = as.factor(CandidateName),
-    State = as.factor(State),
-    Pollster = as.factor(Pollster),
-    StartDate = as.Date(StartDate),
-    EndDate = as.Date(EndDate),
-    Party = as.factor(Party),
-    Methodology = as.factor(Methodology),
-    NumericGrade = as.numeric(NumericGrade),
-    SampleSize = as.numeric(SampleSize)
-  )
-
-# create a new variable called PollRecency
-# PollRecency measures how recently the poll was conducted (based on start date)
-cleaned_data <- cleaned_data %>%
-  mutate(PollRecency = as.numeric(Sys.Date() - StartDate))
-
-# only consider polls for Trump or Harris
-cleaned_data <- cleaned_data %>%
-  filter(CandidateName %in% c("Donald Trump", "Kamala Harris")) %>%
-  select(State, PollScore, SampleSize, NumericGrade, PollRecency, Percentage, CandidateName)
-
-# linear model for Donald Trump's poll percentage based on various predictors
-lm_model_trump <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State,
-  data = cleaned_data %>% filter(CandidateName == "Donald Trump")
-)
-
-# linear model for Kamala Harris' poll percentage based on various predictors
-lm_model_harris <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State,
-  data = cleaned_data %>% filter(CandidateName == "Kamala Harris")
-)
-
-# save modified data for the model as a CSV + Parquet and save models as RDS
-write_csv(cleaned_data, here::here("data/analysis_data/electoral_data.csv"))
[…]

### 2024-11-04 paper/paper.qmd
Updated results section

@@
 title: "Trump vs. Harris: A Data-Driven Forecast modeling the 2024 US Presidential Elections"
-subtitle: "A close race but Trump, trumps Harris"
+subtitle: "Trump trumps Harris to become the next President of the United States"
 author: 
@@
 We utilizes linear regression model deploying polling data and demographic factors to predict both popular vote and Electoral College outcomes.
-The findings indicate a close race between Donald Trump and Kamala Harris, and we predict Trump will secure the Electoral College due to projected victories in key swing states while Harris maintains a popular vote lead.
-The results emphasize the role of state-level dynamics, particularly in swing states, in determining election outcomes."
+The findings indicate a close race between Donald Trump and Kamala Harris, and we predict Trump will secure the Electoral College and Popular Vote due to key wins in tight races for the Battleground States.
+The results emphasize the role of state-level dynamics, particularly in Battleground states, in determining election outcomes."
 format: pdf
@@
 
-The model results show a clear distribution of support for both candidates across various states.
-The predicted percentages of support were normalized to ensure they summed up to 100%.
-After computing these results, we forecast that Donald Trump is likely to win 302 electoral votes, while Kamala Harris is expected to secure 233 electoral votes.
+The national averages for both candidates, as shown in @tbl-popularvote, show that Donald Trump is predicted to receive an average of 45.26% of the popular vote, which, when normalized, represents 50.96% of the total forecasted support, while Kamala Harris is predicted to receive 43.56% with a normalized value of 49.04%.
+In terms of electoral votes, @tbl-electoral shows that Donald Trump is projected to secure 302 votes, while Kamala Harris is expected to gain 233 votes.
+
+The predicted outcomes for each state are summarized in @tbl-statewinners.
+Donald Trump leads in several key battleground states, including Pennsylvania (45.68% for Trump vs.
+44.83% for Harris), Georgia (46.94% vs.
+43.49%), and Michigan (44.86% vs.
+43.87%), while Kamala Harris secures leads in states such as Wisconsin (45.50% for Harris vs.
+45.33% for Trump) and Minnesota (44.73% vs.
+41.64%).
+@fig-predictedscoremap is a visual representation of how these state outcomes appear on an electoral map.
+
+The polling data includes an average poll score of -0.379, with an average sample size of approximately 1,605 individuals per poll.
+The average percentage of support across all states for the candidates is 33.68%.
+This dataset contains a total of 15,829 individual polls, allowing for the model to be trained off a significant number of datapoints.
+The summary statistics are displayed in @tbl-sum.
 
@@
 #| label: tbl-sum
-#| fig-cap: summary statistics for the cleaned data
+#| tbl-cap: summary statistics for the cleaned data
 #| echo: false

### 2024-11-04 paper/paper.qmd
Updated discussion

@@
 NumericGrade: A numerical representation of the reliability of the pollster.
+
+Although not a variable in the model, this variable is used to filter predictions for the model:
 
@@
 
-In this paper, we developed and implemented a predictive model to forecast the outcome of the 2024 U.S.
-Presidential election, focusing on both popular vote percentages and Electoral College results.
-The model incorporates polling data, demographic information, and state-specific effects.
+## What did we do?
 
-The model teaches us that state-level dynamics play a crucial role in determining the winner of U.S.
-Presidential elections.
-Even though Harris is expected to lead in the popular vote, Trump’s success in key battleground states suggests a victory in the Electoral College.
-An additional insight is that poll recency and pollster reliability significantly affect the accuracy of electoral predictions.
-By weighing recent and higher-quality polls more heavily, the model improves the robustness of its forecasts.
+This paper analyzes and predicts the 2024 U.S.
+presidential election outcomes through regression models based on polling data for two main candidates, Donald Trump and Kamala Harris.
+Using a state-by-state approach, it evaluates each candidate’s predicted percentage of support and calculates the anticipated electoral votes to provide an overall electoral projection.
+The model incorporates the predictor factors: PollScore, NumericGrade, SampleSize, State, and PollRecency.
 
-While the model adopts a linear regression framework for its simplicity and interpretability, this choice may limit its ability to capture more complex relationships among variables.
-Future iterations might benefit from exploring nonlinear models or advanced machine learning techniques, which could uncover patterns not readily observable through traditional methods.
+## What did we learn?
 
-One limitation of this model is that it depends entirely on polling data, which may not always capture late shifts in voter preferences or turnout variations.
-Furthermore, the model does not account for external factors like campaign events or last-minute developments, which can significantly influence electoral outcomes.
+This paper reveals the regional polarization in the U.S.
+and illustrates how strong state-based support can determine election outcomes, especially in battleground states.
+From this analysis, we observe that regions such as the Midwest and the South tend to favor conservative candidates, while the Northeast and West Coast are typically more supportive of liberal candidates.
+Trump’s predicted success in key battleground states, such as Pennsylvania and Michigan, leads to a victory in the Electoral College.
+Meanwhile, the difference in the predicted percentages for the popular vote is less than 2%.
+Although the popular vote is predicted to be nearly balanced, with a margin of less than 2% separating the candidates, the narrow margins in key battleground states demonstrate why every vote is so important.
+Small shifts in turnout or voter sentiment can truly change the overall outcome of the election, and thus the future of the United States, dramatically.
+A prime example of this took place in Georgia.
 
-Future research should aim to integrate alternative data sources, such as social media sentiment or economic indicators, to provide a more comprehensive view of voter preferences.
-More advanced machine learning techniques could also be explored to improve the accuracy of future election predictions.
+In 2020, Georgia turned blue in a presidential election for the first time since 1992 [@georgia2020].
+A major factor in this shift was the influx of young, diverse individuals in the Atlanta suburban area.
+In Gwinnett County, the non-Hispanic white population decreased from 67% in 2000 to around 35% by 2020 [@georgiaturnedblue].
+Mitt Romney, a Republican, had won Gwinnett County by 9 points in 2012.
+In 2020, it shifted to an 18-point win for Joe Biden.
+Additionally, there was a major campaign by Stacey Abrams to mobilize previously disengaged voters, particularly people of color [@georgiaturnedblue].
+All of this led to Georgia becoming a swing state after 28 years of voting red.
+Biden received only 0.2% more of the vote than Trump in Georgia.
+
+## What are some weaknesses?
+
+One limitation is its reliance on polling data, which can be inherently biased due to sampling issues, response rates, and potential biases in who chooses to respond to polls.
+Moreover, as the election date approaches, polling data may fail to capture recent shifts in voter sentiment resulting from unforeseen events or major campaign developments.
+For example, last-minute political controversies or debates can alter public opinion quickly.
+
+For instance, a recent incident involving derogatory comments about Puerto Rico made at a Trump rally has rapidly influenced the views of Puerto Rican voters in Pennsylvania.
+Puerto Rican voters, who make up nearly half a million people in the state, have voiced anger and disapproval, with some even reversing their voting plans [@Hill2024].
+WhatsApp groups, morning radio shows, and community discussions have amplified these sentiments, illustrating how quickly backlash can spread through influential networks, particularly in swing states [@Hill2024].
+Our model predicts Trump will win Pennsylvania by a margin of 0.8%, which is less than the 3.7% of Pennsylvania's Population that Puerto Ricans make up [@USCensus2020].
+Situations like these could easily cause Pennsylvania's 19 electoral votes to swing the other way.
+
+Another limitation is the assumption that past trends will remain consistent across elections, which does not account for the ongoing demographic and ideological shifts within states.
+For instance, the influx of highly educated individuals into certain states can shift voting patterns over time, as research indicates that Americans with a college degree or higher are more likely to vote for liberal candidates [@pew2016].
+Should a state’s population of college-educated voters grow significantly, this change could lead to unexpected voting outcomes that our model might not fully capture.
+
+Additionally, the model excludes third-party candidates, who, while unlikely to win, often garner a small but meaningful percentage of votes.
+Independent candidates historically attract around 2-5% of the popular vote, diverting votes from major party candidates, which can be especially impactful in closely contested states.
+This factor may lead to underestimations in one candidate’s projected vote share if independents disproportionately draw support from one side.
+
+These limitations point to a larger issue: that forecasting an election relies on data subject to change and that unexpected events and demographic trends can disrupt otherwise reliable patterns.
+
+## What is left to learn? How should we proceed in the future?
+
+Future research could better the model by incorporating live data updates and adjusting for factors such as voter turnout projections and demographic shifts at the state level.
+For example, as states experience growth in college-educated residents or younger populations, these demographic factors may gradually influence overall voting patterns.
+A model that can adapt ro reflect these trends could improve predictive accuracy over time.
+Additionally, including third-party candidates in the model would provide a more well-rounded view of the vote breakdown, as these candidates can meaningfully impact the final distribution of votes between the two primary candidates.
+As previously mentioned, third-party candidates can capture a small but notable portion of votes, which can tilt tight races in battleground states.
+Future models could potentially account for this by assessing third-party candidates’ historical performance to estimate the effect on major candidates’ vote shares.
 

### 2024-11-04 paper/paper.qmd
Updated predictor variable section

@@
 
-Key predictors include:
+In this analysis, several key predictors were identified to evaluate polling data effectively.
+The first variable, **PollScore**, is a numeric representation of the pollster's credibility, which stands for "Predictive Optimization of Latent Skill Level in Surveys, Considering Overall Record, Empirically." This score helps assess the error and bias associated with a specific pollster, where lower values (including negative numbers) indicate better credibility [@PollsterRating].
 
-PollScore: A numeric value that reflects the credibility of the pollster.
-Stands for "Predictive Optimization of Latent skill Level in Surveys, Considering Overall Record, Empirically." The error and bias we can attribute to a pollster.
-Negative numbers are better.
+Next, **SampleSize** refers to the number of respondents involved in each poll.
+A larger sample size generally enhances the reliability of poll results, as it is more likely to represent the broader population accurately.
+The **State** variable denotes the U.S.
+state where the poll was conducted, in order to give geographical context.
 
-SampleSize: The number of respondents in each poll.
+**PollRecency** is a variable we created to measure the number of days between the start date of the poll and the present.
+This is important as more recent polls tend to reflect the current political climate and voter sentiment more accurately than polls that were conducted a year ago, or more.
+Lastly, **NumericGrade** is a numerical assessment of the pollster's reliability.
 
-State: The U.S. state where the poll was conducted.
-
-PollRecency: The number of days between the start date of the poll and the present, which gives more weight to recent polls.
-
-NumericGrade: A numerical representation of the reliability of the pollster.
-
-Although not a variable in the model, this variable is used to filter predictions for the model:
-
-CandidateName: The name of the candidate being polled (either Donald Trump or Kamala Harris).
+While **CandidateName** is not a predictor variable within the model, it serves as a filter for the predictions being analyzed.
+This allows us to compare between the two primary candidates—Donald Trump and Kamala Harris.
 
@@
 
-$$Percentage_T = \beta_0 + \beta_1 Pollscore_i + \beta_2 SampleSize + \beta_3 Numeric Grade_i + \beta_4 State_i + \beta_5PollRecency + \epsilon_i$$In this model, $Percentage$ represents the predicted support percentage for a candidate in a given poll.
+$$Percentage_T = \beta_0 + \beta_1 Pollscore_i + \beta_2 SampleSize + \beta_3 Numeric Grade_i + \beta_4 State_i + \beta_5PollRecency + \epsilon_i$$
+
+In this model, $Percentage$ represents the predicted support percentage for a candidate in a given poll.
 $Percentage_H$ and $Percentage_T$ represent the predicted support for Harris and Trump respectively.

### 2024-11-04 paper/paper.qmd
described variables in data section

@@
 This variable is crucial for estimating both the popular vote and Electoral College outcomes.
+We see a bimodal distribution in @fig-percentage.
+This indicates that candidates are either recieveing very high percentages of support or very low percentages of support.
+When we pull out Harris and Trump and view their distributions versus all candidates in @fig-candidatepercentage, we see that these two candidates see statistically high percentages of support, indicating they are leading the polls.
 
@@
 The first variable, **PollScore**, is a numeric representation of the pollster's credibility, which stands for "Predictive Optimization of Latent Skill Level in Surveys, Considering Overall Record, Empirically." This score helps assess the error and bias associated with a specific pollster, where lower values (including negative numbers) indicate better credibility [@PollsterRating].
+@fig-pollscoredistribution reveals whether the majority of pollsters tend to have high or low credibility scores.
+We see a higher concentration of negative scores, which suggests a trend toward more reputable pollsters in the dataset.
 
@@
 A larger sample size generally enhances the reliability of poll results, as it is more likely to represent the broader population accurately.
+In @fig-samplesize we see that the vast majority of polls in the dataset have a sample size of less than 2500.
+As discussed in @sec-appx, this number maximizes efficiency and accuracy while minimizing the margin of error.
 The **State** variable denotes the U.S.
 state where the poll was conducted, in order to give geographical context.
+All 50 states are represented in this analysis, and Nebraska and Maine are broken up into electoral districts.
+Washington DC is not included in this dataset.
+@fig-statesample shows the number of polls conducted in each state.
+Battleground states notably have a higher number of polls conducted.
 
@@
 This is important as more recent polls tend to reflect the current political climate and voter sentiment more accurately than polls that were conducted a year ago, or more.
+In @fig-pollrecency we see that the majority of polls were conducted more recently than not, however polls date back until the last presidential election in 2020.
 Lastly, **NumericGrade** is a numerical assessment of the pollster's reliability.
+Our dataset displays polls that tend to have a higher numeric grade as depicted in @fig-numericgrade.
 
@@
 This allows us to compare between the two primary candidates—Donald Trump and Kamala Harris.
+
+@fig-corrmatrix reveals relationships among the key variables in the dataset.
+The strong negative correlation between PollScore and NumericGrade (-0.89) indicates that as the PollScore of the pollster decreases (lower PollScore values are better), the reliability of the pollster increases (higher NumericGrade values are preferable).
+This shows that more credible pollsters are likely to be more reliable, which makes sense given the context.
+
+PollRecency exhibits a weak positive correlation with SampleSize (0.07) and Percentage (0.23), suggesting that more recent polls are slightly associated with a larger number of respondents and may result in higher outcome percentages (which may be due to a reduction in the number of candidates as the election grows closer).
+However, the correlation with PollScore is very weak (-0.05), indicating that the recency of polls does not significantly impact their credibility or reliability.
+
+Percentage, as the outcome variable, shows a weak positive correlation with SampleSize (0.16), suggesting that larger sample sizes may be associated with higher percentages, although the effect is modest.
+There is a slight negative correlation with PollScore (-0.02), reinforcing the notion that higher percentages do not consistently relate to the reliability of the pollsters.
 
@@
 ```{r}
+#| label: fig-numericgrade
+#| fig-cap: The distribution of numeric grades assigned to each poll
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+ggplot(cleaned_data, aes(x = NumericGrade)) +
+  geom_histogram(bins = 30,
+                 fill = "purple",
+                 color = "black",
+                 alpha = 0.7) +
+  labs(title = "Distribution of Numeric Grade",
+       x = "Numeric Grade", 
+       y = "Frequency")+
+  theme_minimal()
+```
+
+```{r}
 #| label: fig-percentage
@@
 #| warning: false
+#| 
+#| 
 #| message: false
@@
 ```{r}
-#| label: fig-statepercentage
-#| fig-cap: The percentage of support seen by state
+#| label: fig-statesample
+#| fig-cap: The number of polls conducted in each state.
 #| echo: false
@@
 #| message: false
-ggplot(cleaned_data, aes(x = State, y = Percentage)) +
-  geom_boxplot() +
-  labs(title = "Percentage by State", x = "State", y = "Percentage") +
+
+poll_counts <- cleaned_data %>%
+  group_by(State) %>%
+  summarise(Count = n())
+
+ggplot(poll_counts, aes(x = State, y = Count)) +
+  geom_bar(stat = "identity") +
+  labs(title = "Number of Polls by State", x = "State", y = "Number of Polls") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 6))
+
 ```
@@
 numerical_data <- cleaned_data %>%
-  select(PollScore, SampleSize, Percentage)
+  select(PollScore, SampleSize, PollRecency, NumericGrade)
 
@@
 ggcorrplot(correlation_matrix, lab = TRUE, title = "Correlation Matrix")
+```
+
+```{r}
+#| label: fig-pollrecency
+#| fig-cap: The distribution of how recent a poll was conducted
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+ggplot(cleaned_data, aes(x = PollRecency)) +
+  geom_histogram(bins = 30,
+                 fill = "pink",
+                 color = "black",
+                 alpha = 0.7) +
+  labs(title = "Distribution of Poll Recency",
+       x = "PollRecency", 
+       y = "Frequency")+
+  theme_minimal()
 ```

### 2024-11-04 scripts/01-test_simulated_data.R
ran styler on scripts

@@ -65,8 +65,9 @@ test_results$date_logic <- test_that("Date logic", {
 
 # test if the election dates are within the valid range (2024-11-01 to 2024-11-08)
 test_results$election_date <- test_that("Election date range", {
-  expect_true(all(data$election_date >= as.Date("2024-11-01") & data$election_date <= as.Date("2024-11-08")), 
-              info = "Some election_dates are out of range (2024-11-01 to 2024-11-08).")
+  expect_true(all(data$election_date >= as.Date("2024-11-01") & data$election_date <= as.Date("2024-11-08")),
+    info = "Some election_dates are out of range (2024-11-01 to 2024-11-08)."
+  )
 })
 
 # test if the candidate parties are valid (only Party A, Party B, Party C are allowed)

### 2024-11-04 scripts/03-clean_data.R

@@ -77,11 +77,14 @@ cleaned_data <- tryCatch(
   }
 )
 
-# save cleaned data as new csv if successful, print error if not
+# add poll recency
+cleaned_data <- cleaned_data %>%
+  mutate(PollRecency = as.numeric(Sys.Date() - StartDate))
+
+# save cleaned data as new csv and parquet if successful, print error if not
 if (!is.null(cleaned_data)) {
   write_csv(cleaned_data, here::here("data/analysis_data/president_polls_cleaned.csv"))
   write_parquet(cleaned_data, here::here("data/analysis_data/president_polls_cleaned.parquet"))
-  
   print(head(cleaned_data))
 } else {
   message("Cleaned data not available for saving due to an error.")

### 2024-11-04 scripts/04-test_analysis_data.R

@@ -22,7 +22,7 @@ test_results <- list()
 # test structure of data
 test_results$structure <- test_that("Data structure", {
   expect_true(is.data.frame(cleaned_data), info = "clean_data is not a data frame.")
-  expect_equal(ncol(cleaned_data), 17, info = "clean_data does not have the correct number of columns.")
+  expect_equal(ncol(cleaned_data), 18, info = "clean_data does not have the correct number of columns.")
   expect_equal(nrow(cleaned_data), 15829, info = "clean_data does not have the correct number of rows.")
 })
 
@@ -33,7 +33,7 @@ test_results$column_names_test <- test_that("Column names", {
     "NumericGrade", "PollScore", "Methodology",
     "State", "StartDate", "EndDate",
     "SampleSize", "Population", "ElectionDate",
-    "Stage", "CandidateName", "Percentage"
+    "Stage", "CandidateName", "Percentage", "PollRecency"
   )
   expect_equal(names(cleaned_data), expected_colnames, info = "Column names do not match expected names.")
 })
@@ -46,29 +46,35 @@ test_results$na_values_test <- test_that("NA values", {
 
 # test if NumericGrade and PollScore values are within expected range
 test_results$numeric_grade_pollscore_test <- test_that("Numeric grade and PollScore values", {
-  expect_true(all(cleaned_data$NumericGrade >= 0 & cleaned_data$NumericGrade <= 3, na.rm = TRUE), 
-              info = "NumericGrade values are out of expected range (0 to 3).")
-  expect_true(all(cleaned_data$PollScore >= -2 & cleaned_data$PollScore <= 2, na.rm = TRUE), 
-              info = "PollScore values are out of expected range (-2 to 2).")
+  expect_true(all(cleaned_data$NumericGrade >= 0 & cleaned_data$NumericGrade <= 3, na.rm = TRUE),
+    info = "NumericGrade values are out of expected range (0 to 3)."
+  )
+  expect_true(all(cleaned_data$PollScore >= -2 & cleaned_data$PollScore <= 2, na.rm = TRUE),
+    info = "PollScore values are out of expected range (-2 to 2)."
+  )
 })
 
 # test if sample size is within expected range
 test_results$sample_size_test <- test_that("Sample size", {
-  expect_true(all(cleaned_data$SampleSize >= 100 & cleaned_data$SampleSize <= 30000, na.rm = TRUE), 
[…]

### 2024-11-04 scripts/05-exploratory_data_analysis.R

@@ -4,7 +4,7 @@
 # Date:
 # Contact:
 # License: MIT
-# Pre-requisites: The `tidyverse`, 'janitor', 'here', 'arrow', and 'reshape2' packages
+# Pre-requisites: The `tidyverse`, 'janitor', 'here', and 'reshape2' packages
 # must be installed
 # Any other information needed? Make sure you are in the `US_Election_2024` rproj
 

### 2024-11-04 scripts/06-model_data.R

@@ -19,18 +19,13 @@ cleaned_data <- cleaned_data %>%
     CandidateName = as.factor(CandidateName),
     State = as.factor(State),
     Pollster = as.factor(Pollster),
-    StartDate = as.Date(StartDate),
-    EndDate = as.Date(EndDate),
     Party = as.factor(Party),
     Methodology = as.factor(Methodology),
     NumericGrade = as.numeric(NumericGrade),
-    SampleSize = as.numeric(SampleSize)
+    SampleSize = as.numeric(SampleSize),
+    PollRecency = as.numeric(PollRecency)
   )
 
-# create a new variable called PollRecency
-# PollRecency measures how recently the poll was conducted (based on start date)
-cleaned_data <- cleaned_data %>%
-  mutate(PollRecency = as.numeric(Sys.Date() - StartDate))
 
 # only consider polls for Trump or Harris
 cleaned_data <- cleaned_data %>%
@@ -39,12 +34,12 @@ cleaned_data <- cleaned_data %>%
 
 # linear model for Donald Trump's poll percentage based on various predictors
 lm_model_trump <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State,
-                     data = cleaned_data %>% filter(CandidateName == "Donald Trump")
+  data = cleaned_data %>% filter(CandidateName == "Donald Trump")
 )
 
 # linear model for Kamala Harris' poll percentage based on various predictors
 lm_model_harris <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State,
-                      data = cleaned_data %>% filter(CandidateName == "Kamala Harris")
+  data = cleaned_data %>% filter(CandidateName == "Kamala Harris")
 )
 
 # save modified data for the model as a CSV + Parquet and save models as RDS

### 2024-11-04 scripts/07-validation.R

@@ -33,17 +33,19 @@ cleaned_data <- cleaned_data %>%
   filter(CandidateName %in% c("Donald Trump", "Kamala Harris"))
 
 # split the data into training (70%) and testing (30%) sets
-set.seed(123)  # Set seed for reproducibility
+set.seed(123) # Set seed for reproducibility
 train_index <- createDataPartition(cleaned_data$CandidateName, p = 0.7, list = FALSE)
 train_data <- cleaned_data[train_index, ]
 test_data <- cleaned_data[-train_index, ]
 
 # fit linear regression models on the training set
 lm_model_trump <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State,
-                     data = train_data %>% filter(CandidateName == "Donald Trump"))
+  data = train_data %>% filter(CandidateName == "Donald Trump")
+)
 
 lm_model_harris <- lm(Percentage ~ PollScore + SampleSize + NumericGrade + PollRecency + State,
-                      data = train_data %>% filter(CandidateName == "Kamala Harris"))
+  data = train_data %>% filter(CandidateName == "Kamala Harris")
+)
 
 # make predictions on the testing set for each model
 predictions_trump <- predict(lm_model_trump, newdata = test_data %>% filter(CandidateName == "Donald Trump"))
@@ -64,4 +66,3 @@ rmse_results <- tibble(
 )
 
 write_parquet(rmse_results, here::here("data/analysis_data/rmse_results.parquet"))
-

### 2024-11-04 paper/paper.qmd
updated data cleaning notes

@@
 
+## Data Cleaning Notes
+
+We began by importing the raw dataset using the read_csv function from the tidyverse package.
+To focus our analysis on more relevant variables, we selected specific columns, such as poll identifiers and polling metrics, omitting any unnecessary columns.
+
+We then renaming columns for clarity. For example, we changed pct to Percentage, making it easier for anyone working with the data to understand what each variable represents.
+
+We converted date data into date format since they were represented as strings in the raw dataset.
+This was done by using the mdy() function, after which we removed any rows with missing values for StartDate, EndDate, or ElectionDate to avoid issues later in our analysis.
+
+We also modified the CandidateName variable to attribute polls for Joe Biden to Kamala Harris, as Joe Biden dropped out of the race and Harris replaced him as the democratic nominee.
+
+We also created a new variable called PollRecency, which indicates how many days have passed since each poll started.
+This was done because as we get closer to election date, polls become more meaningful as they represent the most recent viewpoints of voters.
+
+We also wrapped the cleaning process in a tryCatch block in order to mitigate any errors that arose throughout the cleaning process.
+
+After completing the cleaning, we saved the final dataset in both Parquet and CSV formats for later analysis.
+
 ## Pollster Methodology

### 2024-11-04 paper/paper.qmd
Edited intro

@@
 The objective is to translate these probabilities into a forecast the overall winner of the election.
-Our findings reveal a close race, with Trump holding a slight edge in the Electoral College due to projected victories in key swing states, despite Harris maintaining a lead in national popular vote estimates.
+Our findings reveal a close race, with Trump holding a slight edge in both the popular vote and Electoral College due to projected victories in key swing states.
 

### 2024-11-04 paper/paper.qmd
Updated formatting of paper

@@
 library(kableExtra)
+
+cleaned_data <- read_parquet(here::here("data/analysis_data/president_polls_cleaned.parquet"))
 ```
@@
 
+```{r}
+#| label: fig-percentage
+#| fig-cap: The distribution of what percentage of support a candidate recieves in a poll
+#| echo: false
+#| eval: true
+#| warning: false
+#| 
+#| 
+#| message: false
+ggplot(cleaned_data, aes(x = Percentage)) +
+  geom_histogram(bins = 30,
+                 fill = "orange",
+                 color = "black",
+                 alpha = 0.7) +
+  labs(title = "Distribution of Percentage", 
+       x = "Percentage",
+       y = "Frequency") +
+  theme_minimal()
+```
+
+```{r}
+#| label: fig-candidatepercentage
+#| fig-cap: violin plot displaying the percentage of support Harris and Trump receive compared to all candidates
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+harris_trump_selection <- cleaned_data %>%
+  filter(CandidateName %in% c("Donald Trump", 
+                              "Kamala Harris"))
+
+all_candidates <- cleaned_data %>%
+  mutate(CandidateName = "All Candidates")
+
+combined_data <- bind_rows(
+  harris_trump_selection,
+  all_candidates
+)
+
+ggplot(combined_data, aes(x = CandidateName, 
+                          y = Percentage, 
+                          fill = CandidateName)) +
+  geom_violin(alpha = 0.5) +
+  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) +
+  labs(title = "Percentage by Candidate", 
+       x = "Candidate", 
+       y = "Percentage") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
+  scale_fill_manual(values = c("grey", "salmon", "blue")) +
+  facet_wrap(~CandidateName, scales = "free_x") +
+  theme_minimal()
+```
+
 ## Predictor Variables
@@
 We see a higher concentration of negative scores, which suggests a trend toward more reputable pollsters in the dataset.
+
+```{r}
+#| label: fig-pollscoredistribution
+#| fig-cap: Distribution of Poll Scores by Candidate with denisty overlayed
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+Select_candidates_cleaned_data <- cleaned_data %>%
+  filter(CandidateName %in% c("Donald Trump", "Kamala Harris"))
+
+ggplot(Select_candidates_cleaned_data, aes(x = PollScore)) +
+  geom_histogram(aes(y = ..density..),
+                 bins = 30,
+                 fill = "blue",
+                 color = "black",
+                 alpha = 0.7) +
+    geom_density(color = "black", fill = "salmon", alpha = 0.6) +
+  labs(title = "Distribution of Poll Scores by Candidate",
+       x = "Poll Score", 
+       y = "Frequency") +
+  theme_minimal() +
+  facet_wrap(~ CandidateName)
+```
 
@@
 
+```{r}
+#| label: fig-samplesize
+#| fig-cap: The distribution of sample sizes used in each poll
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+ggplot(cleaned_data, aes(x = SampleSize)) +
+  geom_histogram(bins = 30,
+                 fill = "green",
+                 color = "black",
+                 alpha = 0.7) +
+  labs(title = "Distribution of Sample Size",
+       x = "Sample Size", 
+       y = "Frequency")+
+  theme_minimal()
+```
+
+```{r}
+#| label: fig-statesample
+#| fig-cap: The number of polls conducted in each state.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+poll_counts <- cleaned_data %>%
+  group_by(State) %>%
+  summarise(Count = n())
+
+ggplot(poll_counts, aes(x = State, y = Count)) +
+  geom_bar(stat = "identity") +
+  labs(title = "Number of Polls by State", x = "State", y = "Number of Polls") +
+  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 6))
+```
+
 **PollRecency** is a variable we created to measure the number of days between the start date of the poll and the present.
@@
 Our dataset displays polls that tend to have a higher numeric grade as depicted in @fig-numericgrade.
+
+```{r}
+#| label: fig-pollrecency
+#| fig-cap: The distribution of how recent a poll was conducted
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+ggplot(cleaned_data, aes(x = PollRecency)) +
+  geom_histogram(bins = 30,
+                 fill = "pink",
+                 color = "black",
+                 alpha = 0.7) +
+  labs(title = "Distribution of Poll Recency",
+       x = "PollRecency", 
+       y = "Frequency")+
+  theme_minimal()
+```
+
+```{r}
+#| label: fig-numericgrade
+#| fig-cap: The distribution of numeric grades assigned to each poll
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+ggplot(cleaned_data, aes(x = NumericGrade)) +
+  geom_histogram(bins = 30,
+                 fill = "purple",
+                 color = "black",
+                 alpha = 0.7) +
+  labs(title = "Distribution of Numeric Grade",
+       x = "Numeric Grade", 
+       y = "Frequency")+
+  theme_minimal()
+```
 
@@
 There is a slight negative correlation with PollScore (-0.02), reinforcing the notion that higher percentages do not consistently relate to the reliability of the pollsters.
+
+```{r}
+#| label: fig-corrmatrix
+#| fig-cap: correlation between PollScore, the sample size of the poll, and the percentage of support for a candidate
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+numerical_data <- cleaned_data %>%
+  select(PollScore, SampleSize, PollRecency, NumericGrade)
+
+correlation_matrix <- cor(numerical_data, use = "complete.obs")
+
+library(ggcorrplot)
+ggcorrplot(correlation_matrix, lab = TRUE, title = "Correlation Matrix")
+```
 
@@
 ```
+
+# Results {#sec-result}
+
+The national averages for both candidates, as shown in @tbl-popularvote, show that Donald Trump is predicted to receive an average of 45.26% of the popular vote, which, when normalized, represents 50.96% of the total forecasted support, while Kamala Harris is predicted to receive 43.56% with a normalized value of 49.04%.
+In terms of electoral votes, @tbl-electoral shows that Donald Trump is projected to secure 302 votes, while Kamala Harris is expected to gain 233 votes.
 
@@
 
+The predicted outcomes for states where the margin is 4% or less are summarized in @tbl-statewinners.
+These states are known as our Battleground states (aka swing states).
+Donald Trump leads in several key battleground states, including Pennsylvania (45.68% for Trump vs.
+44.83% for Harris), Georgia (46.94% vs.
+43.49%), and Michigan (44.86% vs.
+43.87%), while Kamala Harris secures leads in states such as Wisconsin (45.50% for Harris vs.
+45.33% for Trump) and Minnesota (44.73% vs.
+41.64%).
+@fig-predictedscoremap is a visual representation of the predicted outcomes for all states and how it would appear as an electoral map.
+
 ```{r}
@@
       TRUE ~ "Tie"
-    )
-  )
-#print(state_winners, n=54)
+    ),
+    Margin = abs(TrumpPredicted - HarrisPredicted)
+  ) %>%
+  filter(Margin <= 4)
 
 state_winners %>%
-  slice(1:53) %>%
+  select(State, TrumpPredicted, HarrisPredicted, Winner) %>%
   kable(
-    caption = "Predicted Percentage of Electoral Votes Per State",
+    caption = "Predicted Percentage of Electoral Votes Per State (Margin ≤ 4%)",
     col.names = c("State", "Trump Predicted %", "Harris Predicted %", "Winner"),
@@
   kable_styling(latex_options = c("striped", "scale_down"))
+
 ```
 
-# Results {#sec-result}
+```{r}
+#| label: fig-predictedscoremap
+#| fig-cap: "Mapped Model-predicted outcomes for 2024 elections by State"
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
 
-The national averages for both candidates, as shown in @tbl-popularvote, show that Donald Trump is predicted to receive an average of 45.26% of the popular vote, which, when normalized, represents 50.96% of the total forecasted support, while Kamala Harris is predicted to receive 43.56% with a normalized value of 49.04%.
-In terms of electoral votes, @tbl-electoral shows that Donald Trump is projected to secure 302 votes, while Kamala Harris is expected to gain 233 votes.
+states_map <- map_data("state")
 
-The predicted outcomes for each state are summarized in @tbl-statewinners.
-Donald Trump leads in several key battleground states, including Pennsylvania (45.68% for Trump vs.
-44.83% for Harris), Georgia (46.94% vs.
-43.49%), and Michigan (44.86% vs.
-43.87%), while Kamala Harris secures leads in states such as Wisconsin (45.50% for Harris vs.
-45.33% for Trump) and Minnesota (44.73% vs.
-41.64%).
-@fig-predictedscoremap is a visual representation of how these state outcomes appear on an electoral map.
+electoral_data <- predictions_with_votes
+electoral_data$State <- tolower(electoral_data$State) # matching state names
+
+merged_data <- inner_join(states_map, electoral_data, by = c("region" = "State"))
+
+map <- ggplot(merged_data, aes(x = long, y = lat, group = group)) +
+  geom_polygon(aes(fill = Winner), color = "white") +
+  scale_fill_manual(values = c("Donald Trump" = "red", "Kamala Harris" = "blue")) +
+  labs(title = "2024 US Presidential Election Predictions by State",
+       fill = "Predicted Winner") +
+  theme_void() +
+  theme(legend.position = "right")
+
+map
+```
 
@@
   kable_styling(latex_options = c("striped", "scale_down"))
-```
-
-```{r}
-#| label: fig-predictedscoremap
-#| fig-cap: "Mapped Model-predicted outcomes for 2024 elections by State"
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-states_map <- map_data("state")
-
-electoral_data <- predictions_with_votes
-electoral_data$State <- tolower(electoral_data$State) # matching state names
-
-merged_data <- inner_join(states_map, electoral_data, by = c("region" = "State"))
-
-map <- ggplot(merged_data, aes(x = long, y = lat, group = group)) +
-  geom_polygon(aes(fill = Winner), color = "white") +
-  scale_fill_manual(values = c("Donald Trump" = "red", "Kamala Harris" = "blue")) +
-  labs(title = "2024 US Presidential Election Predictions by State",
-       fill = "Predicted Winner") +
-  theme_void() +
-  theme(legend.position = "right")
-
-map
-```
-
-```{r}
-#| label: fig-pollscoredistribution
-#| fig-cap: Distribution of Poll Scores by Candidate with denisty overlayed
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-Select_candidates_cleaned_data <- cleaned_data %>%
-  filter(CandidateName %in% c("Donald Trump", "Kamala Harris"))
-
-ggplot(Select_candidates_cleaned_data, aes(x = PollScore)) +
-  geom_histogram(aes(y = ..density..),
-                 bins = 30,
-                 fill = "blue",
-                 color = "black",
-                 alpha = 0.7) +
-    geom_density(color = "black", fill = "salmon", alpha = 0.6) +
-  labs(title = "Distribution of Poll Scores by Candidate",
-       x = "Poll Score", 
-       y = "Frequency") +
-  theme_minimal() +
-  facet_wrap(~ CandidateName)
-```
-
-```{r}
-#| label: fig-samplesize
-#| fig-cap: The distribution of sample sizes used in each poll
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-ggplot(cleaned_data, aes(x = SampleSize)) +
-  geom_histogram(bins = 30,
-                 fill = "green",
-                 color = "black",
-                 alpha = 0.7) +
-  labs(title = "Distribution of Sample Size",
-       x = "Sample Size", 
-       y = "Frequency")+
-  theme_minimal()
-```
-
-```{r}
-#| label: fig-numericgrade
-#| fig-cap: The distribution of numeric grades assigned to each poll
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-ggplot(cleaned_data, aes(x = NumericGrade)) +
-  geom_histogram(bins = 30,
-                 fill = "purple",
-                 color = "black",
-                 alpha = 0.7) +
-  labs(title = "Distribution of Numeric Grade",
-       x = "Numeric Grade", 
-       y = "Frequency")+
-  theme_minimal()
-```
-
-```{r}
-#| label: fig-percentage
-#| fig-cap: The distribution of what percentage of support a candidate recieves in a poll
-#| echo: false
-#| eval: true
-#| warning: false
-#| 
-#| 
-#| message: false
-ggplot(cleaned_data, aes(x = Percentage)) +
-  geom_histogram(bins = 30,
-                 fill = "orange",
-                 color = "black",
-                 alpha = 0.7) +
-  labs(title = "Distribution of Percentage", 
-       x = "Percentage",
-       y = "Frequency") +
-  theme_minimal()
-```
-
-```{r}
-#| label: fig-candidatepercentage
-#| fig-cap: violin plot displaying the percentage of support Harris and Trump receive compared to all candidates
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-harris_trump_selection <- cleaned_data %>%
-  filter(CandidateName %in% c("Donald Trump", 
-                              "Kamala Harris"))
-
-all_candidates <- cleaned_data %>%
-  mutate(CandidateName = "All Candidates")
-
-combined_data <- bind_rows(
-  harris_trump_selection,
-  all_candidates
-)
-
-ggplot(combined_data, aes(x = CandidateName, 
-                          y = Percentage, 
-                          fill = CandidateName)) +
-  geom_violin(alpha = 0.5) +
-  geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) +
-  labs(title = "Percentage by Candidate", 
-       x = "Candidate", 
-       y = "Percentage") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
-  scale_fill_manual(values = c("grey", "salmon", "blue")) +
-  facet_wrap(~CandidateName, scales = "free_x") +
-  theme_minimal()
-```
-
-```{r}
-#| label: fig-statesample
-#| fig-cap: The number of polls conducted in each state.
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-poll_counts <- cleaned_data %>%
-  group_by(State) %>%
-  summarise(Count = n())
-
-ggplot(poll_counts, aes(x = State, y = Count)) +
-  geom_bar(stat = "identity") +
-  labs(title = "Number of Polls by State", x = "State", y = "Number of Polls") +
-  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 6))
-
-```
-
-```{r}
-#| label: fig-corrmatrix
-#| fig-cap: correlation between PollScore, the sample size of the poll, and the percentage of support for a candidate
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-numerical_data <- cleaned_data %>%
-  select(PollScore, SampleSize, PollRecency, NumericGrade)
-
-correlation_matrix <- cor(numerical_data, use = "complete.obs")
-
-library(ggcorrplot)
-ggcorrplot(correlation_matrix, lab = TRUE, title = "Correlation Matrix")
-```
-
-```{r}
-#| label: fig-pollrecency
-#| fig-cap: The distribution of how recent a poll was conducted
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-ggplot(cleaned_data, aes(x = PollRecency)) +
-  geom_histogram(bins = 30,
-                 fill = "pink",
-                 color = "black",
-                 alpha = 0.7) +
-  labs(title = "Distribution of Poll Recency",
-       x = "PollRecency", 
-       y = "Frequency")+
-  theme_minimal()
 ```

### 2024-11-04 paper/paper.qmd
Edited prose

@@
 ---
-title: "Trump vs. Harris: A Data-Driven Forecast modeling the 2024 US Presidential Elections"
+title: "Trump vs. Harris: A 2024 US Presidential Elections Modelling Forecast"
 subtitle: "Trump trumps Harris to become the next President of the United States"
@@
 Republicans and Democrats held primary elections across the country earlier in 2024 that eventually filtered down to the main candidates: Donald Trump and Kamala Harris.
-The polarized political climate and rapidly shifting public opinion have made predicting outcomes increasingly challenging, where winning key swing states is crucial to securing victory.
+The polarized political climate and shifting public opinion have made predicting outcomes increasingly challenging, where winning key swing states is important to securing victory.
 
@@
 The objective is to translate these probabilities into a forecast the overall winner of the election.
-Our findings reveal a close race, with Trump holding a slight edge in both the popular vote and Electoral College due to projected victories in key swing states.
+Our findings show a close race, with Trump holding a slight edge in both the popular vote and Electoral College due to projected victories in key swing states.
 
-Trump's path to victory hinges on critical wins in states such as Pennsylvania, Georgia, and Wisconsin, where his projected margins are narrow but statistically significant.
-Meanwhile, Harris secures strong performances in populous Democratic strongholds like California and New York, but these do not offset Trump’s advantage in less populous yet crucial battleground states.
+Trump's path to victory hinges on vital wins in states such as Pennsylvania, Georgia, and Wisconsin, where his projected margins are narrow but statistically significant.
+Meanwhile, Harris secures strong performances in populous Democratic strongholds like California and New York, but these do not offset Trump’s advantage in less populous yet vital battleground states.
 This result underscores the significant influence of swing states and the Electoral College system in determining the outcome, particularly in closely contested elections.
-These insights contribute to a broader understanding of how statistical modeling can inform political strategy and election analysis.
+This contributes to a broader understanding of how statistical modeling can inform political strategy and election analysis.
 
@@
 Presidential election [@FiveThirtyEightPolls].
-The analysis leverages the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `janitor` [@janitor], `knitr` [@knitr], `dplyr` [@dplyr], `arrow` [@arrow], `purrr` [@purrr], and `here` [@here], for data manipulation.
+The analysis uses the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `janitor` [@janitor], `knitr` [@knitr], `dplyr` [@dplyr], `arrow` [@arrow], `purrr` [@purrr], and `here` [@here], for data manipulation.
 `ggplot2` [@ggplot] and `kableExtra` [@kableExtra] for visualization.
@@
 
-The primary measurement process reflects the transformation of raw polling data into structured entries to actionable insights for election forecasting.
+The primary measurement process reflects the transformation of raw polling data into structured entries to actionable analysis for election forecasting.
 We first select and rename key variables from raw data to focus on relevant information, such as pollster, pollscore, state, polling dates, sample size, and candidate support percentage.
@@
 To address instances where the original data referenced Joe Biden instead of Kamala Harris, we assumed the entries for Biden will goes to Harris in consequences of Democratic Party.
-This approach is supported by "Harris’s truncated presidential campaign relies heavily upon President Biden’s policy framework and she eaddress the campaign’s weaknesses and win over critical voter demographics [@RealInstitutoElcano].
+This approach is supported by "Harris’s truncated presidential campaign relies heavily upon President Biden’s policy framework and she eaddress the campaign’s weaknesses and win over relevant voter demographics [@RealInstitutoElcano].
 Thus, substituting Biden’s polling data with Harris’s ensures continuity in demographic appeal without altering established polling trends.
@@
 The main outcome variable of interest is the Percentage of support each candidate has in a given poll, which represents the proportion of respondents who favor one candidate over the other.
-This variable is crucial for estimating both the popular vote and Electoral College outcomes.
+This variable is needed to estimate both the popular vote and Electoral College outcomes.
 We see a bimodal distribution in @fig-percentage.
@@
 The first variable, **PollScore**, is a numeric representation of the pollster's credibility, which stands for "Predictive Optimization of Latent Skill Level in Surveys, Considering Overall Record, Empirically." This score helps assess the error and bias associated with a specific pollster, where lower values (including negative numbers) indicate better credibility [@PollsterRating].
-@fig-pollscoredistribution reveals whether the majority of pollsters tend to have high or low credibility scores.
+@fig-pollscoredistribution shows whether the majority of pollsters tend to have high or low credibility scores.
 We see a higher concentration of negative scores, which suggests a trend toward more reputable pollsters in the dataset.
@@
 
-@fig-corrmatrix reveals relationships among the key variables in the dataset.
+@fig-corrmatrix displays relationships among the key variables in the dataset.
 The strong negative correlation between PollScore and NumericGrade (-0.89) indicates that as the PollScore of the pollster decreases (lower PollScore values are better), the reliability of the pollster increases (higher NumericGrade values are preferable).
@@
 $Percentage_H$ and $Percentage_T$ represent the predicted support for Harris and Trump respectively.
-Each predictor variable is carefully selected to capture critical factors influencing voter support, with coefficients $\beta_i$ representing the strength of each factor's impact.
+Each predictor variable is carefully selected to capture factors influencing voter support, with coefficients $\beta_i$ representing the strength of each factor's impact.
 The error term $\epsilon_i$ accounts for random variability not explained by the model.
@@
 
-This paper reveals the regional polarization in the U.S.
-and illustrates how strong state-based support can determine election outcomes, especially in battleground states.
+This paper shows how regional polarization in the U.S.
+and strong state-based support can determine election outcomes, especially in battleground states.
 From this analysis, we observe that regions such as the Midwest and the South tend to favor conservative candidates, while the Northeast and West Coast are typically more supportive of liberal candidates.
@@
 
-For instance, a recent incident involving derogatory comments about Puerto Rico made at a Trump rally has rapidly influenced the views of Puerto Rican voters in Pennsylvania.
+For instance, a recent incident involving derogatory comments about Puerto Rico made at a Trump rally has quickly influenced the views of Puerto Rican voters in Pennsylvania.
 Puerto Rican voters, who make up nearly half a million people in the state, have voiced anger and disapproval, with some even reversing their voting plans [@Hill2024].
@@
 However, a survey with limited response options and questions may mean a respondent's view is not captured in its full breadth.
-It also means that more nuanced opinions may be overlooked in favor of broader trends.
+It also means that more granular opinions may be overlooked in favor of broader trends.
 

### 2024-11-04 paper/paper.qmd
Updated paper to include copy of survey

@@
 #| label: fig-candidatepercentage
-#| fig-cap: violin plot displaying the percentage of support Harris and Trump receive compared to all candidates
+#| fig-cap: Violin plot displaying the percentage of support Harris and Trump receive compared to all candidates
 #| echo: false
@@
 #| label: fig-pollrecency
-#| fig-cap: The distribution of how recent a poll was conducted
+#| fig-cap: The distribution of how recently a poll was conducted
 #| echo: false
@@
 #| label: fig-corrmatrix
-#| fig-cap: correlation between PollScore, the sample size of the poll, and the percentage of support for a candidate
+#| fig-cap: The correlation between the numerical predictor variables and the percentage of support for a candidate
 #| echo: false
@@
 #| label: tbl-sum
-#| tbl-cap: summary statistics for the cleaned data
+#| tbl-cap: Summary statistics for the cleaned poll data
 #| echo: false
@@
 
+## Idealized Survey
+
 An example of the survey can be found at this link: <https://forms.gle/igzK683kHLYytGvu8>
+
+**2024 US Presidential Election Forecast Survey**
+
+**Conducted by:** Sophia Brothers, Deyi Kong, and Rayan Awad Alim\
+**Contact:** sophia.brothers\@mail.utoronto.ca
+
+Welcome, and thank you for taking part in the **2024 US Presidential Election Forecast Survey**. Your participation will help us gain insights into voter preferences, influential issues, and factors shaping the election across all 50 states and Washington D.C.
+
+As a thank you for your time, all survey respondents will be automatically entered into a sweepstakes, with a **chance to win \$500**. One winner will be selected from each state and Washington D.C.
+
+**Your responses will remain strictly confidential** and anonymous, and any data collected will be securely stored and used solely for academic and analytical purposes.
+No identifying information will be connected to your survey responses.
+
+------------------------------------------------------------------------
+
+**Section 1: Demographics**
+
+**State of Residence**\
+- \[List of all 50 states and Washington D.C.\]
+
+**Zip Code**
+
+**Age**\
+1. 18-24\
+2. 25-34\
+3. 35-44\
+4. 45-54\
+5. 55-64\
+6. 65+
+
+**Gender**\
+1. Male\
+2. Female\
+3. Non-binary/Other\
+4. Prefer not to say
+
+**Race/Ethnicity**\
+- White\
+- Black or African American\
+- Hispanic or Latino\
+- Asian\
+- Native American or Alaska Native\
+- Native Hawaiian or Pacific Islander\
+- Other\
+- Prefer not to say
+
+**Income Level**\
+- Less than \$25,000\
+- \$25,000 - \$49,999\
+- \$50,000 - \$74,999\
+- \$75,000 - \$99,999\
+- \$100,000 - \$149,999\
+- \$150,000 or more
+
+**Highest Level of Education**\
+- Less than high school\
+- High school diploma or equivalent\
+- Some college, no degree\
+- Associate degree\
+- Bachelor’s degree\
+- Graduate or professional degree
+
+**Party Affiliation**\
+- Republican\
+- Democrat\
+- Independent\
+- Other
+
+------------------------------------------------------------------------
+
+**Section 2: Voter Registration**
+
+**Are you registered to vote in the upcoming election?**\
+- Yes\
+- No
+
+**How likely are you to vote in the upcoming election?**\
+1 (Very Unlikely) to 5 (Very Likely)
+
+------------------------------------------------------------------------
+
+**Section 3: Voting Intent**
+
+**If the election were held today, who would you vote for?**\
+- Kamala Harris\
+- Donald Trump\
+- Other
+
+**How certain are you of your choice?**\
+1 (Not Certain at All) to 5 (Very Certain)
+
+**How likely are you to vote on election day?**\
+1 (Not Likely) to 5 (Very Likely)
+
+------------------------------------------------------------------------
+
+**Section 4: Key Issues**
+
+**Rank the most important issues influencing your vote:**\
+- Economy\
+- Healthcare\
+- Immigration\
+- Climate Change\
+- National Security\
+- Social Issues (e.g., LGBTQ+ rights, abortion)\
+- Education\
+- Taxes\
+- Gun Control\
+- Foreign Policy\
+- Public Safety\
+- Infrastructure
+
+------------------------------------------------------------------------
+
+**Section 5: Candidate Perception**
+
+**Rate the following candidates on Trustworthiness:**\
+- Kamala Harris: Very low, Low, Neutral, High, Very high\
+- Donald Trump: Very low, Low, Neutral, High, Very high
+
+**Rate the following candidates on Competence:**\
+- Kamala Harris: Very low, Low, Neutral, High, Very high\
+- Donald Trump: Very low, Low, Neutral, High, Very high
+
+**Rate the following candidates on Leadership:**\
+- Kamala Harris: Very low, Low, Neutral, High, Very high\
+- Donald Trump: Very low, Low, Neutral, High, Very high
+
+------------------------------------------------------------------------
+
+**Section 6: Prior Voting History**
+
+**Did you vote in the last presidential election?**\
+- Yes\
+- No
+
+**If yes, who did you vote for?**\
+- Joe Biden\
+- Donald Trump\
+- Other
+
+------------------------------------------------------------------------
+
+Thank you for your time and participation!
+Your responses will help us understand the key issues shaping the upcoming election.
+As a reminder, all participants are automatically entered into a sweepstakes for a chance to win \$500, with one winner selected from each state and Washington D.C.
 

### 2024-11-04 paper/paper.qmd
ran through styler

@@
 ggplot(cleaned_data, aes(x = Percentage)) +
-  geom_histogram(bins = 30,
+  geom_histogram(
+    bins = 30,
     fill = "orange",
     color = "black",
-                 alpha = 0.7) +
-  labs(title = "Distribution of Percentage", 
+    alpha = 0.7
+  ) +
+  labs(
+    title = "Distribution of Percentage",
     x = "Percentage",
-       y = "Frequency") +
+    y = "Frequency"
+  ) +
   theme_minimal()
@@
 harris_trump_selection <- cleaned_data %>%
-  filter(CandidateName %in% c("Donald Trump", 
-                              "Kamala Harris"))
+  filter(CandidateName %in% c(
+    "Donald Trump",
+    "Kamala Harris"
+  ))
 
@@
 
-ggplot(combined_data, aes(x = CandidateName, 
+ggplot(combined_data, aes(
+  x = CandidateName,
   y = Percentage,
-                          fill = CandidateName)) +
+  fill = CandidateName
+)) +
   geom_violin(alpha = 0.5) +
   geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) +
-  labs(title = "Percentage by Candidate", 
+  labs(
+    title = "Percentage by Candidate",
     x = "Candidate",
-       y = "Percentage") +
+    y = "Percentage"
+  ) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
@@
     color = "black",
-                 alpha = 0.7) +
+    alpha = 0.7
+  ) +
   geom_density(color = "black", fill = "salmon", alpha = 0.6) +
-  labs(title = "Distribution of Poll Scores by Candidate",
+  labs(
+    title = "Distribution of Poll Scores by Candidate",
     x = "Poll Score",
-       y = "Frequency") +
+    y = "Frequency"
+  ) +
   theme_minimal() +
-  facet_wrap(~ CandidateName)
+  facet_wrap(~CandidateName)
 ```
@@
 ggplot(cleaned_data, aes(x = SampleSize)) +
-  geom_histogram(bins = 30,
+  geom_histogram(
+    bins = 30,
     fill = "green",
     color = "black",
-                 alpha = 0.7) +
-  labs(title = "Distribution of Sample Size",
+    alpha = 0.7
+  ) +
+  labs(
+    title = "Distribution of Sample Size",
     x = "Sample Size",
-       y = "Frequency")+
+    y = "Frequency"
+  ) +
   theme_minimal()
@@
 ggplot(cleaned_data, aes(x = PollRecency)) +
-  geom_histogram(bins = 30,
+  geom_histogram(
+    bins = 30,
     fill = "pink",
     color = "black",
-                 alpha = 0.7) +
-  labs(title = "Distribution of Poll Recency",
+    alpha = 0.7
+  ) +
+  labs(
+    title = "Distribution of Poll Recency",
     x = "PollRecency",
-       y = "Frequency")+
+    y = "Frequency"
+  ) +
   theme_minimal()
@@
 ggplot(cleaned_data, aes(x = NumericGrade)) +
-  geom_histogram(bins = 30,
+  geom_histogram(
+    bins = 30,
     fill = "purple",
     color = "black",
-                 alpha = 0.7) +
-  labs(title = "Distribution of Numeric Grade",
+    alpha = 0.7
+  ) +
+  labs(
+    title = "Distribution of Numeric Grade",
     x = "Numeric Grade",
-       y = "Frequency")+
+    y = "Frequency"
+  ) +
   theme_minimal()
@@
 rmse_results %>%
-  kable(caption = "RMSE Results for Candidate Models",
+  kable(
+    caption = "RMSE Results for Candidate Models",
     col.names = c("Model", "Root Mean Squared Error (RMSE)"),
-        digits = 2) %>%
+    digits = 2
+  ) %>%
   kable_styling(latex_options = c("striped", "scale_down"))
-
 ```
@@
 cleaned_data <- cleaned_data %>%
-  mutate(PredictedTrumpPercentage = predict(lm_model_trump, newdata = cleaned_data),
-         PredictedHarrisPercentage = predict(lm_model_harris, newdata = cleaned_data))
+  mutate(
+    PredictedTrumpPercentage = predict(lm_model_trump, newdata = cleaned_data),
+    PredictedHarrisPercentage = predict(lm_model_harris, newdata = cleaned_data)
+  )
 
@@
   kable_styling(latex_options = c("striped", "scale_down"))
-
 ```
@@
     PollRecency = mean(PollRecency, na.rm = TRUE),
-    .groups = 'drop'
+    .groups = "drop"
   )
@@
 electoral_votes <- tibble(
-  State = c("Alabama", "Kentucky", "North Dakota", "Alaska", "Louisiana", 
+  State = c(
+    "Alabama", "Kentucky", "North Dakota", "Alaska", "Louisiana",
     "Ohio", "Arizona", "Maine", "Oklahoma", "Arkansas",
@@
     "New York", "Wisconsin", "Kansas", "North Carolina",
-            "Wyoming"),
-  Votes = c(9, 8, 3, 3, 8, 
+    "Wyoming"
+  ),
+  Votes = c(
+    9, 8, 3, 3, 8,
     17, 11, 4, 7, 6,
@@
     28, 10, 6, 16,
-            3)
+    3
+  )
 )
@@
   scale_fill_manual(values = c("Donald Trump" = "red", "Kamala Harris" = "blue")) +
-  labs(title = "2024 US Presidential Election Predictions by State",
-       fill = "Predicted Winner") +
+  labs(
+    title = "2024 US Presidential Election Predictions by State",
+    fill = "Predicted Winner"
+  ) +
   theme_void() +
@@
     digits = 3,
-    booktabs= TRUE
+    booktabs = TRUE
   ) %>%

### 2024-11-04 paper/paper.qmd
added comments to visuals

--- 2024-11-04 paper/paper.qmd
@@
 required_packages <- c(
-  "tidyverse", "sf", "here", "arrow", "ggplot2", "janitor", "purrr", "knitr", "kableExtra"
+  "tidyverse", "sf", "here", "arrow", "ggplot2", "janitor", "purrr", "knitr", "kableExtra", "ggcorrplot"
 )
@@
 
+# load libraries
 library(tidyverse)
@@
 library(kableExtra)
+library(ggcorrplot)
 
+# load data
 cleaned_data <- read_parquet(here::here("data/analysis_data/president_polls_cleaned.parquet"))
@@
 #| warning: false
-#| 
-#| 
 #| message: false
+
+# create plot for Distribution of Percentage
 ggplot(cleaned_data, aes(x = Percentage)) +
-  geom_histogram(bins = 30,
+  geom_histogram(
+    bins = 30,
     fill = "orange",
     color = "black",
-                 alpha = 0.7) +
-  labs(title = "Distribution of Percentage", 
+    alpha = 0.7
+  ) +
+  labs(
+    title = "Distribution of Percentage",
     x = "Percentage",
-       y = "Frequency") +
+    y = "Frequency"
+  ) +
   theme_minimal()
@@
 
+# filter for trump/harris
 harris_trump_selection <- cleaned_data %>%
-  filter(CandidateName %in% c("Donald Trump", 
-                              "Kamala Harris"))
+  filter(CandidateName %in% c(
+    "Donald Trump",
+    "Kamala Harris"
+  ))
 
@@
 
+# create data frame with trump vs harris vs all candidates
 combined_data <- bind_rows(
@@
 
-ggplot(combined_data, aes(x = CandidateName, 
+# create plot for Distribution of Percentage by candidate
+ggplot(combined_data, aes(
+  x = CandidateName,
   y = Percentage,
-                          fill = CandidateName)) +
+  fill = CandidateName
+)) +
   geom_violin(alpha = 0.5) +
   geom_boxplot(width = 0.1, color = "black", outlier.shape = NA) +
-  labs(title = "Percentage by Candidate", 
+  labs(
+    title = "Percentage by Candidate",
     x = "Candidate",
-       y = "Percentage") +
+    y = "Percentage"
+  ) +
   theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
@@
 
+# filter for candidate
 Select_candidates_cleaned_data <- cleaned_data %>%
@@
 
+# create plot for Distribution of Poll Scores by candidate
 ggplot(Select_candidates_cleaned_data, aes(x = PollScore)) +
@@
 
+# create plot for distribution of sample sizes by candidate
 ggplot(cleaned_data, aes(x = SampleSize)) +
-  geom_histogram(bins = 30,
+  geom_histogram(
+    bins = 30,
     fill = "green",
     color = "black",
-                 alpha = 0.7) +
-  labs(title = "Distribution of Sample Size",
+    alpha = 0.7
+  ) +
+  labs(
+    title = "Distribution of Sample Size",
     x = "Sample Size",
-       y = "Frequency")+
+    y = "Frequency"
+  ) +
   theme_minimal()
@@
 
+# count polls
 poll_counts <- cleaned_data %>%
@@
 
+# create plot for # of polls per state
 ggplot(poll_counts, aes(x = State, y = Count)) +
@@
 
+# create plot for distribution of poll recency
 ggplot(cleaned_data, aes(x = PollRecency)) +
-  geom_histogram(bins = 30,
+  geom_histogram(
+    bins = 30,
     fill = "pink",
     color = "black",
-                 alpha = 0.7) +
-  labs(title = "Distribution of Poll Recency",
+    alpha = 0.7
+  ) +
+  labs(
+    title = "Distribution of Poll Recency",
     x = "PollRecency",
-       y = "Frequency")+
+    y = "Frequency"
+  ) +
   theme_minimal()
@@
 
+# create plot for distribution of numeric grades
 ggplot(cleaned_data, aes(x = NumericGrade)) +
-  geom_histogram(bins = 30,
+  geom_histogram(
+    bins = 30,
     fill = "purple",
     color = "black",
-                 alpha = 0.7) +
-  labs(title = "Distribution of Numeric Grade",
+    alpha = 0.7
+  ) +
+  labs(
+    title = "Distribution of Numeric Grade",
     x = "Numeric Grade",
-       y = "Frequency")+
+    y = "Frequency"
+  ) +
   theme_minimal()
@@
 #| message: false
+
+# filter to numerical predictor variables
 numerical_data <- cleaned_data %>%
@@
 
-library(ggcorrplot)
+# create correlation matrix
 ggcorrplot(correlation_matrix, lab = TRUE, title = "Correlation Matrix")
@@
 #| message: false
+
+# read data
 rmse_results <- read_parquet(here::here("data/analysis_data/rmse_results.parquet"))
 
+# create table for rmse results per model
 rmse_results %>%
-  kable(caption = "RMSE Results for Candidate Models",
+  kable(
+    caption = "RMSE Results for Candidate Models",
     col.names = c("Model", "Root Mean Squared Error (RMSE)"),
-        digits = 2) %>%
+    digits = 2
+  ) %>%
   kable_styling(latex_options = c("striped", "scale_down"))
-
 ```
@@
 
+# load data and model
 cleaned_data <- read_parquet(here::here("data/analysis_data/electoral_data.parquet"))
@@
 
+# make predictions
 cleaned_data <- cleaned_data %>%
-  mutate(PredictedTrumpPercentage = predict(lm_model_trump, newdata = cleaned_data),
-         PredictedHarrisPercentage = predict(lm_model_harris, newdata = cleaned_data))
+  mutate(
+    PredictedTrumpPercentage = predict(lm_model_trump, newdata = cleaned_data),
+    PredictedHarrisPercentage = predict(lm_model_harris, newdata = cleaned_data)
+  )
 
+# average predictions
 results <- cleaned_data %>%
@@
 
+# normalize percentage
 total_percentage <- sum(results$NationalTrumpAverage, results$NationalHarrisAverage, na.rm = TRUE)
@@
 
+# display avg percent and normalized percent
 results %>%
@@
 
+# load data and models
 cleaned_data <- read_parquet(here::here("data/analysis_data/electoral_data.parquet"))
@@
 
+# list states
 unique_states <- distinct(cleaned_data, State)
 
+# variables for each state
 prediction_data <- cleaned_data %>%
@@
     PollRecency = mean(PollRecency, na.rm = TRUE),
-    .groups = 'drop'
+    .groups = "drop"
   )
-
+# make predictions
 prediction_data <- prediction_data %>%
@@
 
+# assign electoral votes
 electoral_votes <- tibble(
-  State = c("Alabama", "Kentucky", "North Dakota", "Alaska", "Louisiana", 
+  State = c(
+    "Alabama", "Kentucky", "North Dakota", "Alaska", "Louisiana",
     "Ohio", "Arizona", "Maine", "Oklahoma", "Arkansas",
@@
     28, 10, 6, 16,
-            3)
+    3
+  )
 )
 
+# join predictions to state and votes
 predictions_with_votes <- prediction_data %>%
@@
 
+# determine who wins the state
 predictions_with_votes <- predictions_with_votes %>%
@@
 
+# count total electoral votes
 electoral_totals <- predictions_with_votes %>%
@@
 
+# display electoral votes
 electoral_totals %>%
@@
 
+# determine percent breakdown by states where margin is less than or equal to 4
 state_winners <- predictions_with_votes %>%
@@
 
+# display battle ground states margin in table
 state_winners %>%
@@
 
+# map electoral votes to state
 map <- ggplot(merged_data, aes(x = long, y = lat, group = group)) +
@@
 #| message: false
+
+# load data
 cleaned_data <- read_parquet(here::here("data/analysis_data/president_polls_cleaned.parquet"))
 
+# determine summary stats
 summary_stats <- cleaned_data %>%
@@
 
-# print(summary_stats)
-
+# display summary stats
 summary_stats %>%

### 2024-11-04 paper/paper.qmd
added figure to discussion

@@
 Our model predicts Trump will win Pennsylvania by a margin of 0.8%, which is less than the 3.7% of Pennsylvania's Population that Puerto Ricans make up [@USCensus2020].
-Situations like these could easily cause Pennsylvania's 19 electoral votes to swing the other way.
+@fig-swingstates shows specifically the battleground states that are under contention.
+Situations like these could easily cause Pennsylvania's 19 electoral votes to swing the other way, turning from red to blue.
+
+```{r}
+#| label: fig-swingstates
+#| fig-cap: "Mapped Model-predicted outcomes for 2024 elections by Swing State"
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# filter for swing states
+battleground_states <- predictions_with_votes %>%
+  select(State, TrumpPredicted, HarrisPredicted) %>%
+  mutate(
+    Margin = abs(TrumpPredicted - HarrisPredicted),
+    Winner = case_when(
+      TrumpPredicted > HarrisPredicted ~ "Donald Trump",
+      HarrisPredicted > TrumpPredicted ~ "Kamala Harris",
+      TRUE ~ "Tie"
+    )
+  ) %>%
+  filter(Margin <= 4) %>%
+  select(State, Winner)
+
+battleground_states <- battleground_states %>%
+  mutate(State = tolower(State))
+
+merged_data <- states_map %>%
+  left_join(battleground_states, by = c("region" = "State"))
+
+# map that displays electoral college winners for swing states
+map <- ggplot(merged_data, aes(x = long, y = lat, group = group)) +
+  geom_polygon(aes(fill = Winner), color = "white") +
+  scale_fill_manual(values = c("Donald Trump" = "red", "Kamala Harris" = "blue", "NA" = "grey")) +
+  labs(
+    title = "2024 US Presidential Election Battleground States",
+    fill = "Predicted Winner"
+  ) +
+  theme_void() +
+  theme(legend.position = "right")
+
+map
+```
 

### 2024-11-04 paper/paper.qmd
redid measurement section

@@
 
-The primary measurement process reflects the transformation of raw polling data into structured entries to actionable analysis for election forecasting.
-We first select and rename key variables from raw data to focus on relevant information, such as pollster, pollscore, state, polling dates, sample size, and candidate support percentage.
-These variables represent essential poll characteristics that capture public sentiment in response to real-world campaign dynamics and demographic differences across states.
+The dataset used in this analysis originates from various polling organizations, as listed in @PollsterRating, that conducted surveys to gauge public opinion on the presidential candidates, Donald Trump and Kamala Harris.
+
+The measurement process begins with the identification of the target population—likely voters—who are sampled to reflect the broader electorate.
+Polling organizations use various methodologies to gather data, such as telephone interviews, online surveys, and in-person polling.
+Each methodology has it's own implications for accuracy and representativeness.
+
+Polling organizations aggregate responses and apply statistical methods to estimate the percentages of support for each candidate.
+This aggregation process converts individual voter opinions into a summary statistic—the Percentage—that reflects the overall sentiment within the sampled population.
+These become rows in the raw dataset, reflecting the percentage of support for different pollsters, candidates, methodologies, etc.
+
+## Data Cleaning
+
+We first select and rename key variables from raw data to focus on relevant information.
+The key variables of interest in our analysis include PollScore, SampleSize, NumericGrade, PollRecency, State, and Percentage, with PollRecency being a variable we created.
 
@@
 The data was cleaned by removing irrelevant or incomplete observations, and new variables were constructed where necessary.
+
+More information on the data cleaning process can be found in @sec-appx.
 
@@
 
-While other complex models could be appropriate in this instance, these models were chosen for its balance between simplicity and predictive power.
-Linear regression allows us to capture relationships between polling quality, sample size, and recency, while still providing interpretable coefficients.
+We considered both linear regression (LM) and generalized linear models (GLM) to predict support percentages from polling data.
+We chose linear regression for its simplicity, where the coefficients indicate how each predictor affects the response variable directly.
+Linear regression is effective when the response variable is continuous and normally distributed, which fits our data well.
+While GLMs offer more flexibility, they can be more complex and harder to interpret.
+They can also lead to overfitting if you have many predictors.
+
+Linear regression allows us to capture relationships between our predictor variables, while still providing interpretable coefficients.
 Given the nature of polling data, this approach is both appropriate and commonly used in election forecasting.
-A linear model allows us to directly interpret the influence of each predictor on voter support.
 

### 2024-11-04 paper/paper.qmd
updated references

@@
 Presidential election [@FiveThirtyEightPolls].
-The analysis uses the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `janitor` [@janitor], `knitr` [@knitr], `dplyr` [@dplyr], `arrow` [@arrow], `purrr` [@purrr], and `here` [@here], for data manipulation.
-`ggplot2` [@ggplot] and `kableExtra` [@kableExtra] for visualization.
+The analysis uses the statistical programming language R [@citeR] and several libraries, including `tidyverse` [@tidyverse], `janitor` [@janitor], `knitr` [@knitr], `dplyr` [@dplyr], `arrow` [@arrow], `purrr` [@purrr], `sf` [@sf], and `here` [@here] for data manipulation.
+`ggplot2` [@ggplot], `ggcorrplot` [@ggcorrplot] and `kableExtra` [@kableExtra] for visualization.
 The dataset covers various polls conducted across multiple states, capturing the support for each major candidate—Donald Trump and Kamala Harris—along with detailed attributes of the polls.

### 2024-11-06 paper/paper.qmd
updated to reflect suggestions

@@
 required_packages <- c(
-  "tidyverse", "sf", "here", "arrow", "ggplot2", "janitor", "purrr", "knitr", "kableExtra", "ggcorrplot"
+  "tidyverse", "sf", "here", "arrow", "ggplot2", "janitor", "purrr", "knitr", "kableExtra", "ggcorrplot", "modelsummary"
 )
@@
 library(ggcorrplot)
+library(modelsummary)
 
@@
 #| label: fig-percentage
-#| fig-cap: The distribution of what percentage of support a candidate recieves in a poll
+#| fig-cap: This histogram shows the distribution of the percentage of support each candidate receives in various polls.
+The bimodal nature of the distribution suggests that candidates either receive very high or low support, with few receiving moderate levels of backing.
 #| echo: false
@@
   labs(
-    title = "Distribution of Percentage",
     x = "Percentage",
@@
 #| label: fig-candidatepercentage
-#| fig-cap: Violin plot displaying the percentage of support Harris and Trump receive compared to all candidates
+#| fig-cap:  Violin plot comparing the distribution of support between Donald Trump, Kamala Harris, and all other candidates.
+The two candidates exhibit higher-than-average support levels, with Harris and Trump leading the polls in terms of percentage of backing.
 #| echo: false
@@
   labs(
-    title = "Percentage by Candidate",
     x = "Candidate",
@@
 #| label: fig-pollscoredistribution
-#| fig-cap: Distribution of Poll Scores by Candidate with denisty overlayed
+#| fig-cap: This figure shows the distribution of poll scores, which reflect the credibility of pollsters.
+A higher concentration of negative scores suggests that more reputable pollsters dominate the dataset, with credibility generally improving as the score becomes more negative.
 #| echo: false
@@
   labs(
-    title = "Distribution of Poll Scores by Candidate",
     x = "Poll Score",
@@
 #| label: fig-samplesize
-#| fig-cap: The distribution of sample sizes used in each poll
+#| fig-cap: This histogram illustrates the distribution of sample sizes across all polls. The majority of the polls have sample sizes under 2,500.
 #| echo: false
@@
   labs(
-    title = "Distribution of Sample Size",
     x = "Sample Size",
@@
 #| label: fig-statesample
-#| fig-cap: The number of polls conducted in each state.
+#| fig-cap: This bar chart shows the number of polls conducted across U.S. states. Notably, battleground states have a higher number of polls conducted, reflecting their significance in the election.
 #| echo: false
@@
   geom_bar(stat = "identity") +
-  labs(title = "Number of Polls by State", x = "State", y = "Number of Polls") +
+  labs(x = "State", y = "Number of Polls") +
   theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 6))
@@
 #| label: fig-pollrecency
-#| fig-cap: The distribution of how recently a poll was conducted
+#| fig-cap: This histogram shows the distribution of poll recency, indicating that most polls were conducted relatively recently, with fewer polls extending back to the 2020 election cycle.
 #| echo: false
@@
   labs(
-    title = "Distribution of Poll Recency",
     x = "PollRecency",
@@
 #| label: fig-numericgrade
-#| fig-cap: The distribution of numeric grades assigned to each poll
+#| fig-cap: This histogram displays the distribution of numeric grades assigned to each poll, with higher numeric grades reflecting more reliable and accurate polling sources.
 #| echo: false
@@
   labs(
-    title = "Distribution of Numeric Grade",
     x = "Numeric Grade",
@@
 #| label: fig-corrmatrix
-#| fig-cap: The correlation between the numerical predictor variables and the percentage of support for a candidate
+#| fig-cap: This correlation matrix visualizes the relationships between key numerical predictor variables and the outcome variable, Percentage.
+Notable correlations include a strong negative relationship between PollScore and NumericGrade, and a weak positive correlation between PollRecency and Percentage.
 #| echo: false
@@
 # create correlation matrix
-ggcorrplot(correlation_matrix, lab = TRUE, title = "Correlation Matrix")
+ggcorrplot(correlation_matrix, lab = TRUE)
 ```
@@
 
-The model was implemented using R, with out-of-sample testing and Root Mean Square Error (RMSE) calculations applied to assess predictive accuracy as predicted in @tbl-rmse.
+The model was implemented using R, with out-of-sample testing and Root Mean Square Error (RMSE) calculations applied to assess predictive accuracy as predicted in @tbl-keyestimates.
 By splitting the data into training and test sets, we evaluated how well the model performs on unseen data.
@@
 
-```{r}
-#| label: tbl-rmse
-#| tbl-cap: The RMSE for each model
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# read data
-rmse_results <- read_parquet(here::here("data/analysis_data/rmse_results.parquet"))
-
-# create table for rmse results per model
-rmse_results %>%
-  kable(
-    caption = "RMSE Results for Candidate Models",
-    col.names = c("Model", "Root Mean Squared Error (RMSE)"),
-    digits = 2
-  ) %>%
+For **PollScore**, the Trump model has a coefficient of -0.745, while the Harris model shows a coefficient of 0.781, indicating an inverse relationship for Trump and a positive relationship for Harris with the outcome.
+The standard errors for these coefficients are 0.266 and 0.350, respectively.
+
+For **PollRecency**, the coefficients are -0.003 for Trump and -0.006 for Harris, with both models showing very small negative relationships.
+The standard errors are minimal, suggesting these estimates are fairly precise.
+
+The **NumericGrade** variablity has a negative coefficient of -0.124 for Trump and a positive coefficient of 1.458 for Harris.
+This suggests a different relationship between the numeric grade and the outcome for each model, with Harris showing a stronger positive effect.
+
+Other model summary statistics include a sample size of 2117 for the Trump model and 2255 for the Harris model.
+The R² values indicate that the Trump model explains 64.5% of the variance, while the Harris model explains 51.5%, suggesting the Trump model provides a better fit.
+The Adjusted R² values are slightly lower but follow the same trend.
+
+The AIC, BIC, Log-Likelihood, and RMSE values all indicate that the Trump model is preferred over the Harris model in terms of model fit, as the Trump model has lower values for these selected metrics.
+
+The full summary statistics for the Harris and Trump models are displayed in @tbl-modelestimatefull in @sec-appx.
+
+```{r}
+#| label: tbl-keyestimates
+#| tbl-cap: Summary of key model estimates for Donald Trump and Kamala Harris, including coefficients for predictors like Poll Score, Poll Recency, and Numeric Grade, with standard errors for each estimate.
+Model performance statistics, such as sample size, R², and adjusted R², are also displayed.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# load data and model
+lm_model_trump <- readRDS(here::here("models/lm_model_trump.rds"))
+lm_model_harris <- readRDS(here::here("models/lm_model_harris.rds"))
+
+model_list <- list("Trump Model" = lm_model_trump, "Harris Model" = lm_model_harris)
+modelsummary(model_list, 
+             output = "kableExtra", 
+             statistic = "std.error", 
+             coef_map = c("PollScore" = "Poll Score", 
+                          "PollRecency" = "Poll Recency", 
+                          "NumericGrade" = "Numeric Grade",
+                          "SampleSize" = "Sample Size")) %>%
   kable_styling(latex_options = c("striped", "scale_down"))
@@
 #| label: tbl-popularvote
-#| tbl-cap: The percentage of support Harris and Trump receive overall
+#| tbl-cap: Predicted average and normalized national popular vote percentages for Donald Trump and Kamala Harris based on model outcomes.
 #| echo: false
@@
   kable(
-    caption = "Predicted National Popular Vote for Donald Trump and Kamala Harris",
     col.names = c("Trump Average Predicted %", "Trump Normalized %", "Harris Average Predicted %", "Harris Normalized %"),
@@
 #| label: tbl-electoral
-#| tbl-cap: The number of electoral college votes Harris vs Trump recieves
+#| tbl-cap:  Projected electoral vote totals for Donald Trump and Kamala Harris, based on state-by-state predictions derived from the model.
+The table summarizes the total electoral votes each candidate is expected to win.
 #| echo: false
@@
   kable(
-    caption = "Predicted Number of Electoral Votes for Donald Trump and Kamala Harris",
     col.names = c("Trumps Total Electoral Votes", "Harris Total Electoral Votes"),
@@
 #| label: tbl-statewinners
-#| tbl-cap: The winners of each state, which determines who gets the electoral college votes for that state.
+#| tbl-cap: Summary of predicted electoral outcomes in battleground states where the margin of predicted support between Donald Trump and Kamala Harris is 4% or less, showing the predicted percentage of the vote for each candidate and the winner in each state.
 #| echo: false
@@
   kable(
-    caption = "Predicted Percentage of Electoral Votes Per State (Margin ≤ 4%)",
     col.names = c("State", "Trump Predicted %", "Harris Predicted %", "Winner"),
@@
 #| label: fig-predictedscoremap
-#| fig-cap: "Mapped Model-predicted outcomes for 2024 elections by State"
+#| fig-cap: A map illustrating the predicted outcomes of the 2024 U.S.
+presidential election by state, based on the model's projections.
+States are color-coded by the predicted winner, with red representing Donald Trump and blue representing Kamala Harris.
 #| echo: false
@@
   labs(
-    title = "2024 US Presidential Election Predictions by State",
     fill = "Predicted Winner"
@@
 #| label: tbl-sum
-#| tbl-cap: Summary statistics for the cleaned poll data
+#| tbl-cap: Summary statistics of the cleaned polling data, including an average poll score of -0.379, an average sample size of 1,605, and 33.68% average support across 15,829 polls.
 #| echo: false
@@
   kable(
-    caption = "Summary Statistics of Model Results",
     col.names = c("Average Poll Score", "Average Sample Size", "Average Percentage", "Total Polls"),
@@
 
-## What did we do?
+## Predicting 2024 U.S. Election Outcomes Using State-by-State Polling Data
 
@@
 
-## What did we learn?
+## Regional Polarization and the Crucial Role of Battleground States in Election Outcomes
 
@@
 
-## What are some weaknesses?
+## Limitations of the Model: Data Bias, Shifting Voter Demographics, and Excluded Third-Party Candidates
 
@@
 #| label: fig-swingstates
-#| fig-cap: "Mapped Model-predicted outcomes for 2024 elections by Swing State"
+#| fig-cap: A map illustrating the predicted outcomes of the 2024 U.S.
+presidential election in battleground states where the margin of predicted support between Donald Trump and Kamala Harris is 4% or less.
 #| echo: false
@@
   labs(
-    title = "2024 US Presidential Election Battleground States",
     fill = "Predicted Winner"
@@
 
-## What is left to learn? How should we proceed in the future?
+## Next Steps: Enhance Prediction Accuracy with Real-Time Data, Demographic Changes, and Third-Party Voting Trend**s**
 
@@
 After completing the cleaning, we saved the final dataset in both Parquet and CSV formats for later analysis.
+
+## Additional Tables & Figures
+
+```{r}
+#| label: tbl-modelestimatefull
+#| tbl-cap: Full model estimates for the Trump and Harris models, including standard errors for each model parameter.
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+model_list <- list("Trump Model" = lm_model_trump, "Harris Model" = lm_model_harris)
+modelsummary(model_list, 
+             output = "kableExtra", 
+             statistic = "std.error",
+             longtable = TRUE) %>%
+  kable_styling(latex_options = c("striped", "scale_down"))
+```
 

### 2024-11-06 paper/paper.qmd
Updated paper

@@
 thanks: "Code and data are available at: [https://github.com/eeeee-cmd/US_Election/](https://github.com/eeeee-cmd/US_Election/)."
-date: today
+date: November 4, 2024
 date-format: long
